### 2020-01-28 aap/Revision/vr-mcmc-revision.tex

\documentclass[bj]{imsart}
\makeatletter
\def\set@curr@file#1{%
  \begingroup
    \escapechar\m@ne
    \xdef\@curr@file{\expandafter\string\csname #1\endcsname}%
  \endgroup
}
\def\quote@name#1{"\quote@@name#1\@gobble""}
\def\quote@@name#1"{#1\quote@@name}
\def\unquote@name#1{\quote@@name#1\@gobble"}
\makeatother
\usepackage{graphics}

\usepackage{lineno,hyperref}
\usepackage{amsfonts,amssymb}
\usepackage{mathtools}
%\usepackage{mathbbol}
\usepackage{bbm}
\usepackage{multirow,algorithm2e}%
\usepackage{xargs}
%\usepackage{showlabels}

\newcommandx{\CPE}[3][1=]{{\mathsf E}_{#1}\left[\left. #2 \, \right| #3 \right]}
\usepackage[textsize=footnotesize]{todonotes} % pour voir les commentaires

\newcommand{\eric}[1]{\todo[color=purple!10, author=Eric]{#1}}
\newcommand{\erici}[1]{\todo[inline,color=purple!10, author=Eric]{#1}}
\newcommand{\denis}[1]{\todo[color=green!10, author=Denis]{#1}}
\newcommand{\denisi}[1]{\todo[inline,color=green!10, author=Denis]{#1}}
\newcommand{\sergey}[1]{\todo[color=red!10, author=Sergey]{#1}}
\newcommand{\sergeyi}[1]{\todo[inline,color=red!10, author=Sergey]{#1}}

\def\rmd{\mathrm{d}}
\def\rme{\mathrm{e}}
\def\rset{\mathbb{R}}
\def\NtrainPath{T}
\def\TrainSet{\mathcal{T}}
\modulolinenumbers[5]
\newcommand{\proofendsign}{$\Box$}
\newtheorem{thm}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{remark}{Remark}
\newtheorem{cor}{Corollary}
\newtheorem{example}{Example}
\newenvironment{proof}{{\noindent \bf Proof }}
 {{\hspace*{\fill}\proofendsign\par\bigskip}}
\def\TV{\operatorname{TV}}
\newcommand{\jac}[1]{\operatorname{J} _{#1}}

%added command for indicator - SS
\newcommand{\indiacc}[1]{\mathbbm{1}_{\{#1\}}}

\newcommand*{\argmin}{\operatornamewithlimits{arg\,min}}
\newcommand*{\const}{\mathrm{const}}
\usepackage[lofdepth,lotdepth]{subfig}
\usepackage{graphicx}
\newcommand*{\ol}{\overline}

\begin{document}

\begin{frontmatter}
\title{Variance reduction for MCMC methods via martingale representations}
\runtitle{Variance reduction for MCMC}
%\thankstext{T1}{This research was }

\begin{aug}
\author{\snm{D. Belomestny}\thanksref{a,b}\ead[label=e1]{denis.belomestny@uni-due.de}}
\and
\author{\snm{E. Moulines}\thanksref{c, b}\ead[label=e2]{eric.moulines@polytechnique.edu}}
\and
\author{\snm{S. Samsonov}\thanksref{b}\ead[label=e3]{svsamsonov@hse.ru}}
\and
\author{\snm{N. Shagadatov}\thanksref{b}\ead[label=e4]{nshagadatov@edu.hse.ru}}

\runauthor{D. Belomestny et al}

\affiliation{Duisburg-Essen University, National Research University Higher School of Economics,  Ecole Polytechnique  and Duisburg-Essen University}

\address[a]{
Duisburg-Essen University,
Faculty of Mathematics,
D-45127 Essen
Germany }

\address[c]{
Centre de Math\'ematiques Appliqu\'ees, UMR 7641, Ecole Polytechnique,
France
}

\address[b]{
National University Higher School of Economics, Moscow, Russia
}

\end{aug}

\begin{abstract}
In this paper we propose an efficient variance reduction approach for  MCMC algorithms relying on a novel discrete time martingale representation for Markov chains. Our approach is fully non-asymptotic and does not require any type of ergodicity or special product structure of the underlying density.  By rigorously analyzing the  convergence properties of the proposed algorithm, we show that it's complexity is indeed asymptotically smaller than one of the original MCMC algorithm. The numerical performance of the new method is illustrated  in the case of Gaussian mixtures and Bayesian regression models.
\end{abstract}

\begin{keyword}[class=MSC]
\kwd[Primary ]{60G40}
\kwd{60G40}
\kwd[; secondary ]{91G80}
\end{keyword}

\begin{keyword}
\kwd{MCMC}
\kwd{variance reduction}
\kwd{martingale representation}
\end{keyword}

\end{frontmatter}

\section{Introduction}

Monte Carlo integration typically has an error variance of the form
$\sigma^{2}/n,$ where $n$ is a sample size and \(\sigma^2\) is the variance of  integrand. We can make the variance smaller
by using a larger value of $n$.  Alternatively,
 we can  reduce $\sigma^2$ instead of increasing the sample size \(n\). To this end, one can try to construct
a new Monte Carlo experiment with the same expectation as the original one
but with a lower variance $\sigma^2$. Methods to achieve this are known as variance
reduction techniques. Variance reduction plays an important role in
Monte Carlo and Markov Chain Monte Carlo methods. Introduction to many of the variance reduction techniques can be found in \cite{christian1999monte}, \cite{rubinstein2016simulation}, \cite{GobetBook} and \cite{glasserman2013monte}. Recently one witnessed a revival of interest in efficient variance reduction methods  for MCMC algorithms, see for example \cite{dellaportas2012control}, \cite{mira2013zero}, \cite{brosse2018diffusion} and references therein.
\par
Suppose that we wish to compute $\pi(f):=\mathsf{E}\left[f(X)\right]$, where $X$
is a random vector-valued in $\mathcal{X}\subseteq\mathbb{R}^{d}$ with a density $\pi$ and $f:$
$\mathcal{X}\to\mathbb{R}$ with $f\in L^2(\pi)$.
The idea of the so-called control variates variance reduction method
 is to find a cheaply computable random variable $\zeta$ with $\mathsf{E}[\zeta]=0$ and \(\mathsf{E}[\zeta^2]<\infty,\)
such that the variance of the r.v. $f(X)-\zeta$ is small.  The complexity of the problem of constructing classes $Z$ of control variates \(\zeta\) satisfying   $\mathsf{E}[\zeta]=0$ essentially depends on the degree of our knowledge on \(\pi.\)
For example, if \(\pi\) is analytically known and satisfies some regularity conditions, one can apply the well-known technique of  polynomial interpolation to construct control variates enjoying  some optimality properties, see, for example, Section~3.2 in \cite{dimov2008monte}. Alternatively, if an orthonormal system in \(L^2(\pi)\) is analytically available, one can build control variates \(\zeta\) as a linear combination of the corresponding basis functions, see \cite{GobetCV}. Furthermore, if \(\pi\) is known only up to a normalizing constant (which is often the case in Bayesian statistics), one can apply the recent approach of control variates  depending only on the gradient \(\nabla \log \pi\)  using Schr\(\ddot{\text{o}}\)dinger-type Hamiltonian operator in  \cite{mira2013zero} and so-called Stein operator in \cite{brosse2018diffusion}.
In some situations \(\pi\) is not known analytically, but \(X\) can be represented as a function of  simple random variables with known distribution.
Such  situation arises, for example, in the case of functionals of  discretized diffusion processes. In this case a Wiener chaos-type decomposition can be used to construct control variates with nice theoretical properties, see \cite{belomestny2018stratified}.
Note that in order to compare different  variance reduction approaches, one has to analyze their complexity, that is, the number of numerical operations required to achieve a prescribed magnitude of the resulting variance.

The situation becomes much more difficult in the case of MCMC algorithms, where one  has to work with a
Markov chain \(X_p,\) \(p=0,1,2,\ldots,\) whose marginal distribution  converges  to \(\pi\) as time grows. One important class of the variance reduction methods in this case  is based on the so-called Poisson equation for the corresponding  Markov chain. It was observed in Henderson~\cite{henderson1997variance}  that if a time-homogeneous Markov chain \((X_p)\) is stationary with stationary distribution \(\pi,\) then for any real-valued function \(G \in L^1(\pi) \) defined on the state space of the Markov chain \((X_p),\)  the function \(U(x) := G(x)-\mathsf{E}[G(X_{1})|X_0 = x]\) has zero mean with respect to \(\pi\).  The best choice for the function \(G\) corresponds to a solution of the so-called Poisson equation  \(\mathsf{E}[G(X_{1})|X_0 = x]-G(x)=-f(x)+\pi(f)\).  Moreover, it is also related to the minimal asymptotic variance in the corresponding central limit theorem, see \cite{duncan2016variance} and \cite{mira2013zero}.   Although the Poisson equation involves the quantity of interest \(\pi(f)\)  and can not be  solved explicitly in most cases, this idea still can be used to construct some  approximations for the optimal zero-variance control variates. For example,  Henderson~\cite{henderson1997variance} proposed to compute approximations for the solution of the Poisson equation for specific Markov chains with particular emphasis on models arising in stochastic network theory. In \cite{dellaportas2012control} and \cite{brosse2018diffusion}  series-type control variates are introduced and studied for reversible Markov chains. It is assumed in \cite{dellaportas2012control}  that the one-step conditional expectations  can be computed explicitly  for a set of basis functions. The authors in \cite{brosse2018diffusion} proposed another approach tailored to diffusion setting which doesn't require the computation of integrals of basis functions and only involves  applications of the underlying generator.
\par
In this paper we focus on the  Langevin type algorithms which got much attention recently, see \cite{dalalyan2017theoretical,durmus:moulines:2017, MR2353037, MR3861816, MR2977986} and references therein. We propose  a generic variance reduction method for these and other types algorithms, which is purely non-asymptotic and does not require that    the conditional expectations of the corresponding Markov chain can be computed  or that the generator is known analytically. Moreover, we do not need to assume stationarity or/and sampling under the invariant distribution \(\pi.\) We rigorously analyse the convergence of the method and study its complexity. It is shown that  our variance-reduced Langevin algorithm outperforms the standard Langevin algorithms in terms of complexity.
\par
The paper is organized as follows.  In Section~\ref{sec:setup} we set up the problem and introduce some notations. Section~\ref{seq:mart_repr} contains a novel martingale representation and shows how this representation can be used for variance reduction. In Section~\ref{sec:ula_analysis} we analyze the performance of the proposed variance reduction algorithm in the case of Unadjusted Langevin Algorithm (ULA). Section~\ref{sec:coeff} studies the complexity of the variance reduced ULA. Finally, numerical examples are presented in Section~\ref{sec:num}.

\section{Setup}\label{sec:setup}
Let  \(\mathcal{X}\) be a domain in \( \mathbb{R}^d.\)  Our aim is to numerically compute  expectations of the form
\[
\pi(f)=\int_{\mathcal{X}} f(x)\pi(\rmd x),
\]
where \(f:\) \(\mathcal{X}\longrightarrow \mathbb{R}\) and \(\pi\) is a probability measure supported on \(\mathcal{X}.\)
If  the dimension of the space \(\mathcal{X}\) is large and \(\pi(f)\) can not be computed analytically, one can apply Monte Carlo methods. However, in many practical situations  direct sampling from \(\pi\) is impossible and this precludes the use of plain Monte Carlo methods in this case. One popular alternative to Monte Carlo  is Markov Chain Monte Carlo (MCMC), where one is looking for a discrete time  (possibly non-homogeneous) Markov chain   \((X_p)_{p\geq 0}\) such that \(\pi\) is its unique invariant measure. In this paper we study a class of MCMC algorithms with \((X_p)_{p\geq 0}\) satisfying the  the following recurrence relation:
\begin{equation}
\label{eq:chain_gen}
X_{p}=\Phi_{p}(X_{p-1},\xi_{p}),\quad p=1,2,\ldots ,\quad X_{0}=x_0,
\end{equation}
for some i.i.d.  random vectors \(\xi_p\in \mathbb{R}^m\) with distribution \(P_{\xi}\)
and some Borel-measurable
functions $\Phi_{p}\colon\mathcal{X}\times\mathbb{R}^{m}\to\mathcal{X}.$
In fact, this is quite general class of Markov chains (see Theorem~1.3.6 in ~\cite{moulines2018})
and many well-known MCMC algorithms can be represented in form \eqref{eq:chain_gen}.
Let us consider two popular examples.
\begin{example}[Unadjusted Langevin Algorithm]
\label{exam:langevin-algorithm}
Fix a sequence of positive time steps \((\gamma_p)_{p\geq 1}.\) Given a Borel function $\mu\colon\mathbb{R}^{d}\to\mathbb{R}^{d}$,
consider a non-homogeneous
discrete-time Markov chain $(X_{p})_{p\geq0}$ defined by
\begin{equation}\label{eq:chain}
X_{p+1}=X_{p}-\gamma_{p+1}\mu(X_{p})+\sqrt{\gamma_{p+1}}Z_{p+1},\end{equation}
where $\left(Z_{p}\right)_{p\geq1}$ is an i.i.d.\ sequence of $d$-dimensional
standard Gaussian random vectors. If
${\mu=\frac{\nabla U}{2}}$
for some continuously differentiable function $U,$ then Markov chain~\eqref{eq:chain} can be used to approximately sample from the density
\begin{equation}\label{eq:stationary_distr}
\pi(x)= \mathrm{const}\,\rme^{-U(x)}, \quad \mathrm{const}= \left. 1 \middle /{\int_{\mathbb{R}^{d}} \rme^{-U(x)}\, \rmd x,} \right.
\end{equation}
provided that \(\int_{\mathbb{R}^{d}} \rme^{-U(x)}\,\rmd x\) is finite. This method is usually referred to as
Unadjusted Langevin Algorithm (ULA).
In fact the Markov chain~\eqref{eq:chain}
arises as the Euler-Maruyama discretization
of the Langevin diffusion
\[
\rmd Y_t=-\mu(Y_t)\,\rmd t+ \rmd W_t
\]
with nonnegative time steps $(\gamma_p)_{p\ge1}$,
and,  under mild technical conditions, the latter Langevin diffusion admits $\pi$
of~\eqref{eq:stationary_distr}
as its unique invariant distribution; see \cite{dalalyan2017theoretical} and \cite{durmus:moulines:2017}.
\end{example}
\begin{example}[Metropolis-Adjusted Langevin Algorithm]
The Metropolis-Hastings algorithm
associated with a target density \(\pi\) requires the choice of a sequence of conditional densities  \((q_p)_{p\geq 1}\) also called proposal or candidate kernels. The transition from the value  of the Markov chain \(X_p\)  at time \(p\)
and its value at time \(p + 1\) proceeds via the following transition step:

\begin{algorithm}[H]
Given \(X_p=x\)\;
\begin{enumerate}
\item Generate \(Y_p\sim q_p(\cdot|x)\)\;
\item Put
\begin{eqnarray*}
X_{p+1}=
\begin{cases}
Y_p, & \texttt{ with probability } \alpha(x,Y_p),
\\
x, &  \texttt{ with probability } 1-\alpha(x,Y_p),
\end{cases}
\end{eqnarray*}
where
\begin{eqnarray*}
\alpha(x,y)=\min\left\{1,\frac{\pi(y)}{\pi(x)}\frac{q_p(x|y)}{q_p(y|x)}\right\}.
\end{eqnarray*}
\end{enumerate}
\end{algorithm}
Then, as shown in Metropolis et al.~\cite{metropolis1953equation}, this transition is reversible with respect to \(\pi\) and therefore preserves the stationary density \(\pi\). If \(q_p\) have a wide enough
support to eventually reach any region
of the state space \(\mathcal{X}\) with positive mass
under \(\pi\), then this transition is irreducible and $\pi$ is a maximal irreducibility measure \cite{mengersen:tweedie:1996}. The  Metropolis-Adjusted Langevin algorithm (MALA) takes  \eqref{eq:chain} as proposal, that is,
\begin{eqnarray*}
q_p(y|x)=(\gamma_{p+1})^{-d/2}\boldsymbol{\varphi}\Bigl([y-x+\gamma_{p+1}\mu(x)]/\sqrt{\gamma_{p+1}}\Bigr),
\end{eqnarray*}
where
$\boldsymbol{\varphi}(z)=(2\pi)^{-d/2} \exp\{-|z|^2/2\}$,
$z\in\mathbb R^d$, denotes the density of a $d$-dimensional
standard Gaussian random vector.  The MALA algorithms usually provide noticeable speed-ups in convergence for most problems. It is not difficult to see that the MALA chain can be compactly represented in the form
\begin{align*}
X_{p+1} &=X_p+\mathbbm{1}\bigl(U_{p+1}\leq \alpha(X_{p},Y_{p})\bigr)(Y_{p}-X_p),  \,
Y_{p}&=X_p-\gamma_{p+1}\mu(X_p)+\sqrt{\gamma_{p+1}}Z_{p+1},
\end{align*}
where \((U_{p})_{p\geq 1}\) is an i.i.d. sequence of uniformly distributed on \([0,1]\) random variables independent of \((Z_p)_{p\geq 1}.\) Thus we recover \eqref{eq:chain_gen} with  \(\xi_p=(U_p,Z_p)\in \mathbb{R}^{d+1}\) and
\begin{eqnarray*}
\Phi_p(x,(u,z)^\top)=x+\mathbbm{1}\bigl(u\leq \alpha(x,x-\gamma_{p}\mu(x)+\sqrt{\gamma_{p}}z)\bigr)(-\gamma_{p}\mu(x)+\sqrt{\gamma_{p}}z).
\end{eqnarray*}
\end{example}
\begin{example}
  Let \((X_t)_{t\in [0,T]}\) be the unique strong solution to a  SDE of the form:
\begin{eqnarray}
\label{eq:sde-inv}
dX_t=b(X_t)\,dt+\sigma(X_t)\,dW_t,\quad t\geq 0,	
\end{eqnarray}
where \(W\) is a standard \(\mathbb{R}^m\)-valued Brownian motion, \(b:\) \(\mathbb{R}^d\to \mathbb{R}^d\) and \(\sigma:\) \(\mathbb{R}^d\times \mathbb{R}^m \to \mathbb{R}^d\) are locally Lipschitz continuous functions with at most linear growth.
The process \((X_t)_{t\geq 0}\) is a Markov
process and let \(L\) denote its infinitesimal generator defined by
\begin{eqnarray*}
Lg=b^\top \nabla g+\frac{1}{2}\text{Tr}(\sigma^\top D^2g\sigma)
\end{eqnarray*}
for any \(g\in C^2(\mathbb{R}^d).\)
If there exists a continuously twice differentiable Lyapunov function \(V:\) \(\mathbb{R}^d\to \mathbb{R}_{+}\) such that
\begin{eqnarray*}
\sup_{x\in \mathbb{R}^d} LV(x) <\infty,\quad \limsup_{|x|\to \infty} LV(x)<0,
\end{eqnarray*}
then there is an invariant probability measure \(\pi\)  for \(X,\) that is, \(X_t\sim \pi\) for all \(t> 0\) if \(X_0\sim \pi.\) Invariant measures are crucial in the study of the long term behaviour of stochastic differential
systems \eqref{eq:sde-inv}.    Under some additional assumptions,  the invariant
measure \(\pi\) is  ergodic and this property
can be exploited  to compute  the integrals \(\pi(f)\) for \(f\in L^2(\pi)\) by means of ergodic averages. The idea is to replace the diffusion \(X\) by a (simulable) discretization scheme of the form (see e.g. \cite{MR3861816})
\begin{eqnarray*}
\bar X_{n+1}=\bar X_n+\gamma_{n+1} b(\bar X_n)+\sigma(\bar X_n)(W_{\Gamma_{n+1}}-W_{\Gamma_n}), \quad n\geq 0,\quad \bar X_0=X_0,
\end{eqnarray*}
where \(\Gamma_n=\gamma_1+\ldots+\gamma_n\) and \((\gamma_n)_{n\geq 1}\) is a non-increasing sequence of time steps. Then for a function \(f\in L^2(\pi)\) we can approximate \(\pi(f)\) via
\begin{eqnarray*}
\pi_n^\gamma(f)=\frac{1}{\Gamma_n}\sum_{i=1}^n \gamma_{i}f(\bar X_{i-1}).
\end{eqnarray*}
Due to typically high correlation between \(X_0,X_1,\ldots\)  variance reduction is of crucial importance here.
As a matter of fact, in many cases there is no explicit formula for the invariant measure and this makes the use of gradient  based  variance reduction techniques (see e.g. \cite{mira2013zero})   impossible in this case.
On the contrary, our method can be directly used to reduce the variance of the ergodic estimator \(\pi_n^\gamma\) without explicit knowledge of \(\pi.\)
\end{example}
\section{Martingale representation and variance reduction}
\label{seq:mart_repr}
In this section we give a general discrete-time martingale representation for the Markov chain \eqref{eq:chain_gen} which is  used below to construct an efficient variance reduction algorithm. Let \((\phi_k)_{k\in \mathbb{Z}_+}\) be a complete orthonormal system in \(L^2(\mathbb{R}^m, P_{\xi})\) with \(\phi_0\equiv 1\).  In particular, we have
\begin{eqnarray*}
\mathsf{E}[\phi_i(\xi)\phi_j(\xi)]=\delta_{ij},\quad i,j\in  \mathbb{Z}_{+}
\end{eqnarray*}
with \(\xi \sim P_{\xi}.\)
Notice that this implies that the random variables
$\phi_k(\xi)$, $k\ge1$, are centered. As an example, we can take  multivariate Hermite polynomials for the ULA algorithm and a tensor product of Shifted Legendre polynomials for "uniform part" and Hermite polynomials for "gaussian part"  of the random variable $\xi = (u, z)^T$ in MALA, as the Shifted Legendre polynomials are orthogonal with respect to the Lebesgue measure on \([0,1].\) In the sequel, we denote by $(\mathcal{G}_p)_{p\in \mathbb{Z}_+}$  the filtration
generated by generated by $(\xi_p)_{p \in \mathbb{N}^*}$ with the convention $\mathcal{G}_0=\mathrm{triv}$.

\begin{thm}\label{prop:29032018a1}
 Let $f$ be a Borel function $\mathbb{R}^{d}\to\mathbb R$ such that
$\mathsf{E}\left[\left|f(X_{p})\right|^{2}\right]<\infty$, with a Markov chain $(X_{p})_{p\geq 0}$ defined in \eqref{eq:chain_gen}.
Then,
for $p>j$, the following representation holds in \(L^2(P)\)
\begin{eqnarray}
\label{eq:mart_repr}
f(X_{p})=\mathsf{E}\left[\left.f(X_{p})\right|\mathcal G_{j}\right]+\sum_{k=1}^{\infty}\sum_{l=j+1}^{p}a_{p,l,k}(X_{l-1})\phi_k\left(\xi_{l}\right),
\end{eqnarray}
where
\begin{eqnarray}
\label{eq:coeff_mart}
a_{p,l,k}(x)=\mathsf{E}\left[\left.f(X_{p})\phi_k\left(\xi_{l}\right)\right|X_{l-1}=x\right], \quad p\geq l, \quad k\in \mathbb{N}.
\end{eqnarray}
\end{thm}
\begin{proof}
The expansion obviously holds for $p=1$ and $j=0$.
Indeed, due to the orthonormality and completeness
of the system $\left(\phi_{k}\right)$, we have
\[
f(X_{1})=\mathsf{E}\left[f(X_{1})\right]+\sum_{k\geq1}a_{1,1,k}(X_0)\phi_{k}(\xi_{1})
\]
with
\[
a_{1,1,k}(x_0)=\CPE{f(X_{1})\phi_{k}\left(\xi_{1}\right)}{X_0=x_0},
\]
provided $\mathsf{E}\left[\left|f(X_{1})\right|^{2}\right]<\infty.$
Recall that $\mathcal{G}_l=\sigma(\xi_{1},\ldots,\xi_{l}),$
$l=1,2,\ldots, $
and $\mathcal{G}_0=\mathrm{triv}$.
Suppose that (\ref{eq:mart_repr})
holds for $p=q$, all $j<q$, and all Borel-measurable functions
$f$ with $\mathsf{E}\left[|f(X_{q})|^2\right]<\infty$.
Let us prove it for $p=q+1$.
Given $f$ with $\mathsf{E}\left[|f(X_{p})|^2\right]<\infty$,
due to the orthonormality and completeness
of the system $\left(\phi_{k}\right)$, we get by
conditioning on $\mathcal{G}_{q}$,
\[
f(X_{p})=\mathsf{E}\left[\left.f(X_{p})\right|{\mathcal{G}}_q\right]+\sum_{k\geq1}\alpha_{p,q+1,k}\phi_{k}(\xi_{q+1}),
\]
where
\begin{equation*}
\alpha_{p,q+1,k}
=\mathsf{E}\left[\left.f(X_{p})\phi_{k}(\xi_{q+1})\right|\mathcal G_q\right].
\end{equation*}
By the Markov property of $(X_{l})$,
we have
$\mathsf{E}[f(X_{p})|\mathcal{G}_q]
=\mathsf{E}[f(X_{p})|X_{q}]$.
Furthermore, a calculation involving
intermediate conditioning on $\mathcal G_{q+1}$
and the recurrence relation
$X_{q+1}=\Phi_{q+1}(X_{q},\xi_{q+1})$
verifies that
\[
\alpha_{p,q+1,k}=\mathsf{E}\left[\left.f(X_{p})\phi_{k}(\xi_{q+1})\right|X_{q}\right]
=a_{p,q+1,k}(X_{q})
\]
for suitably chosen
Borel-measurable functions $a_{p,q+1,k}$.
We thus arrive at
\begin{align}
\label{berm:sig_X}
f(X_{p})=\mathsf{E}\left[\left.f(X_{p})\right|X_{q}\right]+\sum_{k\geq1}a_{p,q+1,k}(X_{q})\phi_{k}(\xi_{q+1}),
\end{align}
which is the required statement in the case $j=q$.
Now assume $j<q$.
The random variable
$\mathsf{E}\left[\left.f(X_{p})\right|X_{q}\right]$
is square integrable and has the form
$g(X_{q})$,
hence the induction hypothesis applies, and we get
\begin{equation}\label{eq:28082017a1}
\mathsf{E}\left[\left.f(X_{p})\right|X_{q}\right]=\mathsf{E}\left[\left.f(X_{p})\right|X_{j}\right]+\sum_{k\geq1}\sum_{l=j+1}^{q}a_{p,l,k}(X_{l-1})\phi_{k}(\xi_{l})
\end{equation}
with
\begin{align*}
a_{p,l,k}(X_{l-1}) &= \mathsf{E}\left[\left.\mathsf{E}\left[\left.f(X_{p})\right|\mathcal{G}_{q}\right]\phi_{k}(\xi_{l})\right|\mathcal{G}_{l-1}\right]
= \mathsf{E}\left[\left.f(X_{p})\phi_{k}(\xi_{l})\right|\mathcal{G}_{l-1}\right]\\
&=\mathsf{E}\left[\left.f(X_{p})\phi_{k}(\xi_{l})\right|X_{l-1}\right].
\end{align*}
Formulas \eqref{berm:sig_X}
and~\eqref{eq:28082017a1} conclude the proof.
\end{proof}
From  numerical point of view another representation of the coefficients \(a_{p,l,k}\)  turns out to be more useful.
\begin{prop}
The coefficients \(a_{p,l,k}\) in \eqref{eq:coeff_mart}  can be alternatively represented as
\begin{eqnarray*}
a_{p,l,k}(x)=\mathsf{E}\left[\phi_k\left(\xi\right)Q_{p,l}\left(\Phi_l(x,\xi)\right)\right]
\end{eqnarray*}
with \(Q_{p,l}(x)=\mathsf{E}\left[\left.f(X_{p})\right|X_{l}=x\right],\) \(p\geq l.\)
%where   a r.v. \(\xi\) independent of \(\mathcal{G}_{l-1}.\)
The functions \((Q_{p,l})_{l=0}^p\)  can be computed by the backward recurrence: $Q_{p,p}(x)=f(x)$ and for $l \in \{0,\cdots,p-1\}$
\begin{eqnarray}
\label{eq:qpl}
Q_{p,l}(x)=\mathsf{E}\left[\left.Q_{p,l+1}(X_{l+1})\right|X_{l}=x\right].
%=\mathsf{E}[Q_{p,l+1}(\Phi_{l+1}(x,\xi))]
\end{eqnarray}
\end{prop}
\par
Next we show how  the representation \eqref{eq:mart_repr} can be used to efficiently reduce the variance of MCMC algorithms.  Let $(\gamma_{p})_{p\in\mathbb N}$ be a sequence of positive and non-increasing
step sizes
with $\sum_{p=1}^\infty \gamma_p=\infty$
and, for $n,l\in\mathbb{N}$, $n\le l$, we set
\[
\Gamma_{n,l}=\sum_{p=n}^{l}\gamma_{p}.
\]
 Consider a weighted average estimator $\pi_{n}^{N}(f)$ of the form
\begin{equation}\label{eq:29032018a2}
\pi_{n}^{N}(f)=\sum_{p=N+1}^{N+n}\omega_{p,n}^{N}f(X_{p}),\quad\omega_{p,n}^{N}=\gamma_{p+1}\Gamma_{N+2,N+n+1}^{-1},
\end{equation}
where $N\in\mathbb N_0$ is the length of the burn-in period and $n\in\mathbb N$
the number of effective samples.
Given $N$ and $n$ as above, for $K\in\mathbb N$, denote
\begin{equation}
M_{K,n}^N(f) =\sum_{k=1}^{K}\sum_{l=N+1}^{N+n}\bar a_{l,k}(X_{l-1})\phi_k(\xi_{l}),
\label{eq:29032018a5}
\end{equation}
where
\begin{align}
\label{eq:first-expression-bar-a-l-k}
\bar{a}_{l,k}(x)
=\sum_{p=l}^{N+n}\omega_{p,n}^{N}a_{p,l,k}(x)=\mathsf{E}\left[\left.\left(\sum_{p=l}^{N+n}\omega_{p,n}^{N}f(X_{p})\right)\phi_k(\xi_{l})\right|X_{l-1}=x\right].
\end{align}
Since \(X_{l-1}\) is independent of \(\xi_{l}\) and \(\mathsf{E}[\phi_k(\xi_{l})]=0,\) \(k\geq 1,\)  the r.v.  \(M_{K,n}^N(f)\) has zero mean and can be viewed as a control variate.
The representation \eqref{eq:coeff_mart} suggests that the variance of  the  variance-reduced estimator
\begin{equation}
\label{eq:29032018a3}
\pi_{K,n}^N(f)=\pi_n^N(f)-M_{K,n}^N(f)
\end{equation}
should be small for \(K\) large enough. In this section   we provide  an analysis of the variance-reduced ULA  algorithm (see Example~\ref{exam:langevin-algorithm}).
\section{Analysis of variance reduced ULA}
\label{sec:ula_analysis}
In this section we perform the convergence analysis of the ULA algorithm. For the sake of simplicity we restrict our attention to the constant time step.
We shall use the notations $\mathbb N=\{1,2,\ldots\}$ and $\mathbb N_0=\mathbb N\cup\{0\}$.
By $H_k$, $k\in\mathbb N_0$,
we denote the normalized Hermite polynomial on $\mathbb R$, that is,
$$
H_k(x)=\frac{(-1)^k}{\sqrt{k!}}\rme^{x^2/2}\frac{\partial^k}{\partial x^k}\rme^{-x^2/2},
\quad x\in\mathbb R.
$$
For a multi-index $\mathbf{k}=(k_i)\in\mathbb N_0^d$,
$\mathbf{H}_\mathbf{k}$ denotes the normalized Hermite polynomial on $\mathbb R^d$, that is,
$$
\mathbf{H}_\mathbf{k}(\mathbf{x})=\prod_{i=1}^d H_{k_i}(x_i),\quad \mathbf{x}=(x_i)\in\mathbb R^d.
$$
In what follows, we also use the notation
$|\mathbf{k}|=\sum_{i=1}^d k_i$ for $\mathbf{k}=(k_i)\in\mathbb N_0^d$,
and we set $\mathcal G_p=\sigma(Z_1,\ldots,Z_p)$, $p\in\mathbb N$, and $\mathcal G_0=\mathrm{triv}$.
Given $N$ and $n$ as above, for $K\in\mathbb N$, denote
\begin{align}
M_{K,n}^N(f) & =\sum_{\mathbf k\colon 0<\|\mathbf{k}\|\le K}\sum_{p=N+1}^{N+n}\omega_{p,n}^{N}\sum_{l=N+1}^{p}a_{p,l,\mathbf{k}}(X_{l-1})\mathbf{H}_\mathbf{k}(Z_{l})
\label{eq:29032018a5}\\
&=\sum_{\mathbf k\colon 0<\|\mathbf{k}\|\le K}\sum_{l=N+1}^{N+n}\bar{a}_{l,\mathbf{k}}(X_{l-1})\mathbf{H}_\mathbf{k}(Z_{l})
\notag
\end{align}
with \(\|\mathbf{k}\|=\max_{i} k_i\) and
\begin{align*}
\bar{a}_{l,\mathbf{k}}(x)
& =\sum_{p=l}^{N+n}\omega_{p,n}^{N}a_{p,l,\mathbf{k}}(x) =\mathsf{E}\left[\left.\left(\sum_{p=l}^{N+n}\omega_{p,n}^{N}f(X_{p})\right)\mathbf{H}_\mathbf{k}\left(Z_{l}\right)\right|X_{l-1}=x\right].
\end{align*}
For an estimator $\rho(f)\in\{\pi_n^N(f),\pi_{K,n}^N(f)\}$
of $\pi(f)$ (see \eqref{eq:29032018a2} and \eqref{eq:29032018a3}), we shall be interested in the Mean Squared Error (MSE),
which can be decomposed as the sum of the squared  bias
and the  variance:
\begin{eqnarray}\label{eq:29032018a4}
\mathrm{MSE}\left[\rho(f)\right]
&=&\mathsf{E}\left[(\rho(f)-\pi(f))^2\right]
\\
\nonumber
&=&\left(\mathsf{E}[\rho(f)]-\pi(f)\right)^2
+\mathsf{Var}[\rho(f)].
\end{eqnarray}
Our analysis will be done under the following two assumptions.

\begin{description}
\item[(H1)][\textbf{Lipschitz continuity}] Assume that $\nabla U$ is Lipschitz, that is, there exists $L_U\geq 0$ such that 
\begin{eqnarray*}
|\nabla U(x)-\nabla U(y)|\leq L_U|x-y|, \quad x,y\in \mathbb{R}^d.
\end{eqnarray*}
\item[(H2)][\textbf{Convexity outside a ball}] There exist $K_U>0,$ $M_U>0$ and $m_U>0$ such that for any $x\not\in B_{K_U}(0)	$ and any $y\in \mathbb{R}^d$ it holds 
\begin{eqnarray*}
\langle D^2U(x)y,y\rangle \geq m_U\|y\|^2, \quad \|D^3 U(y)\| \leq M_U. 
\end{eqnarray*}
\end{description}
First we analyse the biases of the estimators
$\pi_n^N(f)$ and $\pi_{K,n}^N(f).$
\subsection{Squared bias}
Due to the martingale transform structure
of $M_{K,n}^N(f)$,
we have
\[
\mathsf E\left[M_{K,n}^N(f)\right]=0,
\]
Hence both estimators
$\pi_n^N(f)$ and $\pi_{K,n}^N(f)$ have the same conditional bias.
Notice that this remains true also when we substitute
the coefficients $a_{p,l,\mathbf{k}}$ in~\eqref{eq:29032018a5}
with some independent approximations $\widehat a_{p,l,\mathbf{k}}.$ For a bounded Borel function~$f$,
we can estimate the conditional bias
similarly to the beginning of ~\cite[Section~4]{durmus:moulines:2017}:
\begin{align}
\left(\mathsf{E}[\pi_{K,n}^N(f)]-\pi(f)\right)^2
&=\left(\mathsf{E}[\pi_{n}^N(f)]-\pi(f)\right)^2
\label{eq:06042018a2}\\
&\le
\mathrm{osc}(f)^2
\sum_{p=N+1}^{N+n}
\omega_{p,n}^N\,
\|\delta_x Q_\gamma^{p}-\pi(\cdot)\|_{TV}^2,
\notag
\end{align}
where
$\mathrm{osc}(f):=\sup_{x\in\mathbb R^d}f(x)-\inf_{x\in\mathbb R^d}f(x)$,
$\|\mu-\nu\|_{\TV}$ denotes the total variation distance
between probability measures $\mu$ and $\nu$, that is,
$$
\|\mu-\nu\|_{TV}=\sup_{A\in\mathcal B(\mathbb R^d)}
|\mu(A)-\nu(A)|,
$$
$\pi(\cdot)$ denotes the probability measure on $\mathbb R^d$
with density $\pi$ of~\eqref{eq:stationary_distr};
for $\gamma>0$, the Markov kernel $R_\gamma$
is defined by
$$
\mathcal{R}_\gamma(x,A)=\int_{A} \frac{1}{(4 \pi \gamma)^{d/2}} \text{ exp} \left\{ -\frac{1}{4\gamma} \left\| y - x + \gamma \mu(x)\right\|^2\right\}dy,
$$
while, for $k,l\in\mathbb N$, $k\le l$,
the kernel $\mathcal{Q}_{\boldsymbol{\gamma}}^{k,l}$ is
$
\mathcal{Q}_{\boldsymbol{\gamma}}^{k,l}=\mathcal{R}_{\gamma_l}\cdots \mathcal{R}_{\gamma_k},
$
which, finally, provides the  measure
$\mathcal{Q}_{\boldsymbol{\gamma}}^{p}=\mathcal{Q}_{\boldsymbol{\gamma}}^{p,1}$
used in the right-hand side of~\eqref{eq:06042018a2}.
Different specific upper bounds for the squared bias
can be deduced from~\eqref{eq:06042018a2}
using results of Section~3 in~\cite{durmus:moulines:2017}
on bounds in the total variation distance. For the case of a constant time step $\gamma_1=\ldots=\gamma_p=\gamma$ it is known that under assumption {\bf (H1)} the corresponding Markov chain has a unique stationary distribution $\pi_\gamma$, which is different from $\pi$. Moreover, we shall show in Section that  under {\bf (H1)} and {\bf (H2)}
\begin{eqnarray}
\label{eq:bias}
\|\delta_x\mathcal{Q}_{\boldsymbol{\gamma}}^{n}-\pi_\gamma\|_{TV} \leq C\rho^{\gamma n}\left(V(x) + \pi_\gamma(V)\right)
\end{eqnarray}
for  $V(x) = 1 + \|x\|^2,$ some $\rho\in (0,1)$ independent of $\gamma$ and an absolute constant $C>0.$
\subsection{Conditional variance}
An upper bound for the variance of the classical estimator~\eqref{eq:29032018a2}
under {\bf (H1)} and {\bf (H2)} is derived in Section.  In particular, if $\sum_{j=1}^\infty \gamma_j=\infty$ and $\sum_{j=1}^\infty \gamma^2_j<\infty,$ then for any bounded function $f,$
\begin{eqnarray}
\label{eq:var-mc}
\mathsf{Var}[\pi_n^N(f)]\lesssim 1/\Gamma_{N+2,N+n+1},
\end{eqnarray}
where $\lesssim$ stands for inequality up to a constant depending on constants $L_U,$ $m_U,$  $M_U$ and $\|f\|_\infty.$ For a constant time step,  we derive from \eqref{eq:var-mc} that
$\mathsf{Var}[\pi_n^N(f)]\lesssim 1/(n\gamma).$

As for the estimator~\eqref{eq:29032018a3}, 
it follows from~\eqref{eq:29032018a5}, Theorem~\ref{prop:29032018a1} and independence condition for the sequence $(Z_p)_{p \geq 1}$.
\begin{prop}
\label{cond_var_dec}
	For variance-reduced estimator~\eqref{eq:29032018a3}, defined for the Markov Chain~\eqref{eq:chain_gen} the following representation holds
\begin{equation}\label{eq:29032018a6}
\mathsf{Var}\left[\pi_{K,n}^{N}(f)\right]
%=\sum_{\min_i (k_i)\ge K+1}
=\sum_{\mathbf k\colon \|\mathbf k\|\ge K+1}
\sum_{l=N+1}^{N+n}
\mathsf{E}\left[\left|\bar a_{l,\mathbf{k}}(X_{l-1})\right|^{2}\right].
\end{equation}
\end{prop}
One of the main results of this paper is the following  upper bound for the right-hand sides of~\eqref{eq:29032018a6} in the case of ULA algorithm.
\begin{thm}\label{th:mr}
Assume {\bf (H1)} and {\bf (H2)}.
Suppose that $f$  is two times continuously differentiable
and it holds
\begin{eqnarray*}
|\partial^{\mathbf{k}} f(x)|\le  B_f, \quad
x\in\mathbb R^d
\end{eqnarray*}
for all \(\mathbf{k}\) with \(0<\|\mathbf{k}\|\leq  2.\) Let $(\gamma_k)_{k\in\mathbb N}$ be a sequence of positive
and non-increasing step sizes with $\sum_{k=1}^\infty \gamma_k=\infty$ and $\sum_{k=1}^\infty \gamma^2_k<\infty.$ 
Then it holds under 
\begin{eqnarray}
\label{eq:var-bound}
\mathsf{Var}\left[\pi_{K,n}^{N}(f)\right]\lesssim \frac{1}{\Gamma^2_{N+2,N+n+1}},
\end{eqnarray}
where \(\lesssim\) stands for inequality up to a constant not depending  on \(n\) and \(N.\)
\end{thm}
\begin{cor}

\end{cor}

\section{Nonparametric regression}
The functions \((\bar a_{l,k})\) need to be estimated before one can apply the proposed variance reduction approach. This amounts to solve a nonparametric regression problem.
We present a generic regression algorithm. Algorithm starts with estimating the functions $\bar{Q}_{l}$ for $l=N+1,\ldots,N+n,$ where
\begin{equation*}
\bar{Q}_{l}(x)=\sum_{p=l}^{N+n} \omega_{p,n}^{N} Q_{p,l}\left(x\right)=\omega_{l,n}^{N}f(x)+\mathsf{E}\left[\left.\sum_{p=l+1}^{N+n}\omega_{p,n}^{N}f(X_{p})\right|X_{l}=x\right] .
\end{equation*}
We first generate \(\NtrainPath\)  paths conditionally independent of the $\sigma$-algebra generated by the burn-in sequence $X_1,\dots,X_N$:
\[
\TrainSet=\Bigl\{(X^{(s)}_{N+1},\ldots,X^{(s)}_{N+n}),\quad s=1,\ldots,\NtrainPath\Bigr\}
\]
of the chain \(X\)
(the so-called ``training paths'').
The regression algorithm proceeds with estimating the  functions \((\bar Q_{l})\) by solving  the least squares optimization problems
\begin{equation}\label{eq:06042018a1}
{\widehat  Q}_{l}=\argmin_{\psi\in \Psi}\sum_{s=1}^{\NtrainPath} \left | \sum_{p=l}^{N+n}\omega_{p,n}^{N}f(X^{(s)}_{p})-\psi(X^{(s)}_{l})\right|^2
\end{equation}
for \(l=N+1,\ldots, N+n-1.\)
where \(\Psi\) is a  class of  functions on \(\mathbb{R}^d.\)
Next  we estimate the coefficients \( \bar a_{l,k}\) using
the formula
\begin{equation}
\label{eq:a_est_int}
\widehat a_{l,k}(x)=\mathsf{E}\left[\left.\phi_k\left(\xi\right)\widehat Q_{l}\left(\Phi_l(x,\xi)\right)\right | \TrainSet\right] \,,
\end{equation}
where $\xi$ is a random variable independent from $\mathcal{T}$ with law $P_\xi,$
$\Phi_l$ is defined in \eqref{eq:chain_gen} and the expectation is taken according to known distribution \(P_\xi\). Conditioning on $\TrainSet$ underlines that estimation of $\widehat Q_{l}$ derived from training trajectories. In many cases integration can be done in closed analytical form.

 For example, in the case of ULA algorithm we have \(P_\xi$ is the multivariate normal distribution $ \mathcal{N}(0,I_d)\) and \(\Phi_l(x,\xi)=x-\gamma_{l}\mu(x)+\sqrt{\gamma_{l}}\xi.\) Therefore if we use polynomials to approximate \((\bar Q_{l}),\) then \((\widehat a_{l,k})\) can be even computed in closed form; implementations details are provided in  Section~\ref{sec:num}.

Upon estimating  the coefficients \((\widehat a_{l,k})\), one can construct an empirical estimate  of \(M_{K,n}^N(f)\) in the form
\begin{eqnarray*}
\widehat M_{K,n}^N(f) =\sum_{1\leq  k \le K}\sum_{l=N+1}^{N+n}\widehat a_{l,k}(X_{l-1})\phi_k(\xi_{l}).
\end{eqnarray*}
Obviously \(\mathsf{E}[\widehat M_{K,n}^N(f)|\TrainSet]=0\) and \(\widehat M_{K,n}^N(f)\) is indeed a valid control variate in a sense that it
does not introduce any bias.

\section{Complexity analysis for ULA}
\label{sec:coeff}

The following result quantifies the error of estimating the functions \((\bar Q_l)\)
via the algorithm and its proof follows from Theorem~2.2 in \cite{audibert2011robust}.
\begin{thm}\label{berm:theorem:regression_cv}
 Suppose that for any \(l\in \{N+1,\ldots,N+n+1\},\)
\[
\mathsf{E}\left[\left(\sum_{p=l}^{N+n}\omega_{p,n}^{N}f(X_{p})-\bar Q_l(X_l)\right)^4\right]\leq \sigma_l^4,
\]
for some finite positive numbers \(\sigma_{N+1},\ldots, \sigma_{N+n}. \)
Furthermore, assume that \(\Psi=\mathrm{span}\{\psi_1,\ldots,\psi_D\},\) where the functions  \(\psi_1,\ldots,\psi_D\) are uniformly bounded and satisfy
\[
\max_{l}\sup_{g\in \Psi\setminus \{0\}} \|g\|^2_\infty/\mathsf{E}[g^2(X_l) | \mathcal{G}_N]\leq B<\infty.
\]
Then for any values of \(\varepsilon\) and \(\NtrainPath\) such that \(2/\NtrainPath\leq \varepsilon\leq 1\) and
\begin{eqnarray*}
\NtrainPath\gtrsim B^2\left[BD+\log(2/\varepsilon)+\frac{B^2D^2}{\NtrainPath}\right]
\end{eqnarray*}
it holds with probability at least \(1-\varepsilon,\)
\begin{multline}
\mathsf{E}\left[ \left.\Bigl|\bar Q_l(X_l)
-\widehat Q_l(X_l)\Bigr|^2\right | \TrainSet \vee \mathcal{G}_N\right]
\lesssim  \sigma^2_l B
\left(\frac{BD+\log(2/\varepsilon)}{\NtrainPath}+\frac{B^2D^2}{\NtrainPath^2}\right)
\\
 +
\inf_{\psi\in \Psi}
\mathsf{E}\left[\left.\left|\bar Q_l(X_{l})-\psi(X_{l})\right|^{2}\right | \mathcal{G}_N \right], \label{eq:regr_error}
\end{multline}
for \(l=N+1,\ldots,N+n,
\) with \(\lesssim\) standing for inequality up to a universal multiplicative constant.
\end{thm}
Now we are able to give  a bound for the difference between \( M_{K,n}^N\) and \(\widehat M_{K,n}^N.\)
\begin{prop}
\label{cor:dif_m}
Under  conditions of Theorem~\ref{berm:theorem:regression_cv}, we have with probability at least \(1-\varepsilon,\)
\begin{align}
\label{eq:main_bound}
&\mathsf{E}\left[\left.\left|\widehat{M}_{K,n}^{N}(f)-M_{K,n}^{N}(f)\right|^{2}\right | \TrainSet,\mathcal{G}_N\right] \\
\nonumber
&\phantom{\mathsf{E}\widehat{M}_{K,n}^{N}}  \lesssim K\left[\sum_{l=N+1}^{N+n} \sigma^2_l \right]
B
\left(\frac{BD+\log(2/\varepsilon)}{\NtrainPath}+\frac{B^2D^2}{\NtrainPath^2}\right)
\\ \nonumber
& \phantom{\mathsf{E}|\widehat{M}_{K,n}^{N}} + K \sum_{l=n+1}^{N+n}\inf_{\psi\in \Psi}
\mathsf{E}\left[\left.\left|\bar Q_l(X_{l})-\psi(X_{l})\right|^{2}\right | \mathcal{G}_N \right].
\end{align}
\end{prop}
\begin{proof}
Using the conditional Cauchy-Schwarz inequality and  orthonormality of \((\phi_k)_{k\geq 0}\), we derive
\begin{eqnarray*}
\mathsf{E}\left[\left.\left|\widehat{a}_{l,k}(X_{l-1})-\bar a_{l,k}(X_{l-1})\right|^{2}\right | \TrainSet \vee \mathcal{G}_N\right]\leq\mathsf{E}\left[\left.\left|\widehat{Q}_{l}\left(X_{l}\right)-\bar Q_{l}\left(X_{l}\right)\right|^{2}\right | \TrainSet \vee \mathcal{G}_N\right]
\end{eqnarray*}
By the Jensen inequality and orthonormality of \((\phi_k)_{k\geq 0},\)
\begin{multline*}
\mathsf{E}\left[\left.\left|\widehat{M}_{K,n}^{N}(f)-M_{K,n}^{N}(f)\right|^{2}\right | \TrainSet \vee \mathcal{G}_N\right]
\\
\leq \sum_{1\leq k\le K}\sum_{l=N+1}^{N+n}\mathsf{E}\left[\left.\left|\widehat{a}_{l,k}(X_{l-1})-\bar a_{l,k}(X_{l-1})\right|^{2}\right | \TrainSet \vee \mathcal{G}_N\right].
\end{multline*}
\end{proof}
\subsection{Complexity analysis of variance reduction for the ULA algorithm}
In the case of ULA one can bound the quantities \(\sigma_l^2\) using  \(L^p\)-type Sobolev inequalities  (see \cite{Adamczak}), see Remark~\ref{rem:p-norm} in Section~\ref{sec:proofs}.
In particular, we can derive that  under conditions of Theorem~\ref{th:mr} with \(K=1,\)
\begin{eqnarray*}
\sigma_l^2\lesssim \frac{1}{\Gamma^2_{N+2,N+n+1}},\quad l=N+1,\ldots,N+n+1,
\end{eqnarray*}
where \(\lesssim \) stands for the inequality up to a multiplicative constant not depending on \(n\) and \(N.\) Using this inequality and combining \eqref{eq:main_bound} with Theorem~\ref{th:mr}, we get for the variance of \(\widehat \pi_{K,n}^N(f)=\pi_n^N(f)-\widehat M_{K,n}^N(f),\) with probability at least \(1-\varepsilon,\)
\begin{eqnarray}
\nonumber
\mathsf{Var}\left[\left.\widehat \pi_{K,n}^N(f)\right | \TrainSet, \mathcal{G}_N\right]&\lesssim & \frac{n KB}{\Gamma^2_{N+2,N+n+1}}
\left(\frac{BD+\log(2/\varepsilon)}{\NtrainPath}+\frac{B^2D^2}{\NtrainPath^2}\right)
\\
\nonumber
&& + \frac{1}{\Gamma^2_{N+2,N+n+1}}\sum_{l=N+1}^{N+n}\sum_{I\subseteq\{1,\ldots,d\},\, I\neq \emptyset}
\left(\frac{\gamma_{l}}{2}\right)^{|I|K}
\\
\label{eq:ula_red_var}
&& +
K \sum_{l=n+1}^{N+n}\inf_{\psi\in \Psi}
\mathsf{E}\left[\left.\left|\bar Q_l(X_{l})-\psi(X_{l})\right|^{2}\right | \TrainSet \vee \mathcal{G}_N \right].
\end{eqnarray}
In order to assess the complexity of the proposed algorithm,  we first prove that under some conditions the coefficients \(a_{p,l,\mathbf{k}}\) decay exponentially fast as \(|p-l|\to \infty.\)
\begin{lem}
\label{a_decay}
Suppose that \(f,\mu\in C^1(\mathbb{R}^d),\) \(0<b_\mu I\leq\jac{\mu}(x)\leq B_\mu I\) and \(B_\mu\gamma_{l}\leq 1,\) then
\begin{eqnarray*}
\|a_{p,l,\mathbf{k}}\|_{\infty}\leq \sqrt{\gamma_l}\, \|\nabla f\|_\infty\exp\left(-b_{\mu}\sum_{r=l+1}^p \gamma_{r}  \right),\quad \mathbf{k}\in \mathbb{N}^d\setminus \{\mathbf{0}\}.
\end{eqnarray*}
\end{lem}
\begin{cor}
Assume that
\begin{eqnarray}
\label{eq:gamk-decay}
\gamma_k=\gamma_1 k^{-\alpha},\quad \alpha\in (0,1),
\end{eqnarray}
 then
\begin{eqnarray*}
\|a_{p,l,\mathbf{k}}\|_{\infty}\leq \sqrt{\gamma_1} l^{-\alpha/2}\, \|\nabla f\|_\infty\exp\left(-c b_{\mu}\gamma_1 (p^{1-\alpha}-l^{1-\alpha}) \right)
\end{eqnarray*}
for some constant \(c>0.\)
\end{cor}
Suppose now that the chain is close to stationarity, then \(Q_{p,l}\) are functions of \(p-l\) only.
Furthermore,  according to Lemma~\ref{a_decay}, \(a_{p,l,\mathbf{k}}\) are exponentially small for \(p-l\) large.
As a result we only need to compute a logarithmic (in \(n\)) number of functions \(Q_{p,l}.\)
Hence the cost of computing the estimates  \(\widehat {a}_{p,l,\mathbf{k}}(x)\) for \(l=N+1,\ldots,N+n,\) and \(\|\mathbf{k}\|\leq K\) using regression on \(\Psi,\) is of order
\[
\log^{1/(1-\alpha)} (n) K^d  \NtrainPath D^2.
\]
Suppose for simplicity that all functions \(\bar{Q}_l\) are in \(\Psi\) for some \(D>0,\) that is, the third term in \eqref{eq:ula_red_var} is zero.  Then under \eqref{eq:gamk-decay} it is enough to take \(K=\lceil 1/\alpha \rceil+1\) to get
\begin{eqnarray*}
\sum_{l=N+1}^{N+n}\sum_{I\subseteq\{1,\ldots,d\},\, I\neq \emptyset}
\left(\frac{\gamma_{l}}{2}\right)^{|I|K}\leq C
\end{eqnarray*}
for some constant \(C\) not depending on \(d.\) Then we have with high probability
\begin{eqnarray*}
\mathsf{Var}\left[\left.\widehat \pi_{K,n}^N(f)\right | \TrainSet \vee \mathcal{G}_N\right]\lesssim \frac{1}{n^{2(1-\alpha)}}\left[1+\frac{n}{\NtrainPath}\right],
\end{eqnarray*}
with corresponding cost proportional to \(\NtrainPath\log^{1/(1-\alpha)} (n) ,\) provided \(N/n=o(1).\) We should compare this to the standard weighted estimator \(\pi_{n}^N(f)\) with variance
\begin{eqnarray*}
\mathsf{Var}\left[\left. \pi_{n}^N(f)\right |  \mathcal{G}_N\right]\lesssim \frac{1}{n^{1-\alpha}}
\end{eqnarray*}
and cost of order \(n.\) Thus while the cost of achieving
\begin{eqnarray*}
\mathsf{Var}\left[\left. \pi_{n}^N(f)\right | \TrainSet \vee \mathcal{G}_N\right]\leq \varepsilon^2
\end{eqnarray*}
is of order  \(\varepsilon^{-2/(1-\alpha)},\) we get the same bound for \(\pi_{K,n}^N(f)\) at a cost of order
\[
\varepsilon^{-1/(1-\alpha)}\log^{1/(1-\alpha)} (\varepsilon).
\]
In the presence of approximation errors we have to increase \(D\) with \(n\) to balance  the statistical and approximation errors resulting in a smaller complexity reduction. Note that our approach does not require the computation of \(\nabla \log(\pi)\) for the construction of control variates. This leads to an additional complexity reduction especially in statistical application where the computation of \(\nabla \log(\pi)\) can be very costly.
\section{Numerical results}
\label{sec:num}
If each function  \(Q_{p,l}(x) \) can be well approximated by polynomials, then we can use polynomial basis functions in   \eqref{eq:06042018a1} to approximate \(\bar Q_l\). Suppose that we constructed a polynomial approximation  for each function \(\bar Q_{l}\)  in the form:
\begin{equation*}
\widehat{Q}_{l}(x) = \sum_{\|\mathbf{s}\|\leq m} \beta_{\mathbf{s}} x^{\mathbf{s}},\quad s=(s_1,\ldots,s_d)
\end{equation*}
with some coefficients \(\beta_{\mathbf{s}}\in \mathbb{R}.\)
Then using the identity
\[
\xi^j = j! \sum_{r = 0}^{\lfloor j/2 \rfloor} \frac{1}{2^r  r! \sqrt{(j-2r)!}} H_{j-2r}(\xi),\quad \xi \in \mathbb{R},
\]
we derive for all $x \in \mathbb{R}^d$,
\begin{eqnarray*}
\widehat a_{l,\mathbf{k}} (x) &=&\CPE{\mathbf{H}_\mathbf{k}(x) \widehat{Q}_{l}(x - \gamma_l \mu(x) + \sqrt{\gamma_l}\xi)}{\TrainSet \vee \mathcal{G}_N}
= \sum_{\|\mathbf{s}\|\leq m} \beta_{\mathbf{s}} \prod_{i=1}^d P_{i,k_i,s_i}(x),
\end{eqnarray*}
where for all integers $i,k,s_i$ and $x \in \rset^d$,
\begin{align*}
P_{l,k,s}(x) &= \mathsf{E} \left[ H_{k}(\xi_l) (x_l - \gamma_l \mu_l(x) + \sqrt{\gamma_l} \xi_l)^{s} \right]
\end{align*}
is a one-dimensional polynomial (in \(x\)) of degree at most \(s\) with analytically known coefficients. Alternatively, on practice we can work directly with functions \(Q_{p,l}.\) Due to (almost) time-homogenous Markov chain generated by ULA, each function $Q_{p,l}$ can be approximated by function, which depends only on the  $p-l$, that is,
\[
 Q_{p,l}(x) = \mathsf{E} \left[f(X_p) | X_l = x\right] \approx Q^\circ_{p-l}(x)
 \]
Thus we have
\[
  Q^\circ_{r}(x) = \mathsf{E} \left[f(X_{l+r}) | X_l = x \right].
 \]
Consequently, the functions $Q^\circ_r$ can be estimated using a modified least squares criteria
\begin{eqnarray}
\label{eq:Qcircr}
\widehat{Q}^\circ_r = \argmin_{\psi \in \Psi} \sum_{s = 1}^{T} \sum_{l = N + 1}^{N+n-r} \left| f(X_{l+r}^{(s)}) - \psi(X_l^{(s)})\right|^2
\end{eqnarray}
for \(1 \leq r \leq n-1,\) where  $\widehat{Q}^\circ_0(x) = f(x)$ by definition.
Due to Lemma~\ref{a_decay}, it is enough to estimate $Q^\circ_{r}$ for $|r| < n_{\mathrm{trunc}}$ for some truncation level \(n_{\mathrm{trunc}}\) depending on \(d\) and \(\gamma.\) It allows us to use a smaller amount of training trajectories to approximate conditional expectations $Q^\circ_{r}$.
Finally  we construct a truncated version of our estimator:
\[
\pi_{K,n,n_{\mathrm{trunc}}}^N(f) = \pi_n^N(f) -  M_{K,n, n_{\mathrm{trunc}}}^{N}(f),
\]
where
\begin{eqnarray*}
{\widehat  M}_{K,n, n_{\mathrm{trunc}}}^N &=& \sum_{0<\left\| \mathbf{k}\right\| \leq K} \sum_{p=N+1}^{N+n} \omega_{p,n}^N
\\
&&\times
\sum_{l=N+1}^p {\widehat  a}_{p-l,\mathbf{k}}(X_{l-1})\mathbf{H}_{\mathbf{k}}(Z_l) \mathbbm{1}\{ |p-l| < n_{\mathrm{trunc}} \}
\end{eqnarray*}
and
\begin{eqnarray*}
\widehat a_{p-l,\mathbf{k}} (x) &=&\CPE{\mathbf{H}_\mathbf{k}(x) \widehat{Q}^\circ_{p-l}(x - \gamma_l \mu(x) + \sqrt{\gamma_l}\xi)}{\TrainSet \vee \mathcal{G}_N}.
\end{eqnarray*}
\subsection{Gaussian mixtures}
 We consider a sample generated by ULA with $\pi$  given by the mixture of two Gaussian distributions  with equal weights:
\[
\pi (x) = \frac{1}{2(2\pi)^{d/2}} \left( \rme^{\frac{-\| x-a\|_2^2}{2}}  + \rme^{\frac{-\| x+a\|_2^2}{2}} \right), \quad x \in\mathbb R^d
\]
where $a \in \mathbb{R}^d$ is a given vector. The function $U(x)$ and its gradient are given by
\[
U(x) = \frac{1}{2} \|x - a\|_2^2 - \text{log}(1 + \rme^{-2x^\top a}), \,\,
\nabla U(x) = x-a +2a(1 + \rme^{2 x^\top a})^{-1},
\]
respectively.
In our experiments we considered dimensions $d = 2$ and $d = 8$ and take $a=( (2d)^{-1/2}, \dots, (2d)^{-1/2})^\top$.
In order to approximate the expectation \(\pi(f)\) with \(f(x)=\sum_{i=1}^d x_i,\)
 we have used constant step sizes $\gamma_i = \gamma=0.2$ and sampled $n = 10^4$ samples for $d = 2$ and $n = 2 \times 10^3$ for $d = 8$.  We generated $\NtrainPath = 10$ independent "training" trajectories and solved the least squares  problems \eqref{eq:Qcircr} using the first order polynomial approximations for the coefficients \(a_{p,\mathbf{k}}\) as described in the previous section. The truncation level \(n_{\mathrm{trunc}}\) is chosen to be \(50.\) To test our variance reduction algorithm, we generated 100 independent trajectories.
 In Figure ~\ref{fig:1} we compare our approach to variance reduction methods of \cite{mira2013zero} and \cite{brosse2018diffusion}.

 \begin{figure}[tbh]
\centering
\subfloat[Subfigure 1 list of figures text][d = 2]{
\includegraphics[width=0.45\textwidth]{../pictures/GM_2d_f1.png}
\label{fig:subfig1}}
\qquad
\subfloat[Subfigure 2 list of figures text][d = 8]{
\includegraphics[width=0.45\textwidth]{../pictures/GM_8d_f1.png}
\label{fig:subfig2}}
\caption{Boxplots of ergodic averages  from the variance reduced ULA algorithms for the Gaussian mixture model. The compared estimators are the ordinary empirical average  (O), our estimator of control variates (CV-B), zero variance estimator (ZV) and the control variates using diffusion approximation (CV), along with the region obtained by the ordinary ULA of the length $10^2 \times n$ (green regions). \label{fig:1}}
\end{figure}

\subsection{Binary Logistic Regression}

Second experiment considers the problem of logistic regression. Suppose we have i.i.d. sample $\left\{ (\mathbf{X}_i, Y_i)\right\}$ for $ i =1, \dots , m$ with features $\mathbf{X}_i \in \mathbb{R}^d $ and binary labels $Y_i \in \left\{0,1 \right\}$. The binary logistic regression model defines the conditional distribution of $Y$ given $X$ by a logistic function
\begin{eqnarray*}
r(\theta, x) = \frac{e^{\theta^T x}}{1 + e^{\theta^T x}},
\end{eqnarray*}
where $\theta$ is a parameter of model. In order to estimate $\theta$ according to given data, the Bayessian approach introduces prior distribution $\pi_0(\theta)$ and inferences the posterior density $\pi(\theta)$ using Bayes' rule.

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.6\textwidth]{../pictures/BLR_banknotes_4d.png}
\caption{Boxplots of ergodic averages  from the variance reduced ULA algorithms for the Logistic Regression. The compared estimators are the ordinary empirical average  (O), our estimator of control variates (CV-B), zero variance estimator (ZV) and control variates using diffusion approximation (CV).
 \label{fig:2}}
\end{center}
\end{figure}
In the case of  Gaussian prior $\pi_0$ with zero mean and covariance matrix $\sigma^2 I_d$, the posterior density takes the form:
\begin{eqnarray*}
\pi(\theta) \propto \exp \left\{ -  \mathbf{Y}^T \mathbf{X \theta} - \sum_{i=1}^m \text{log} (1 + e^{-\mathbf{\theta}^T \mathbf{X}_i}) - \frac{1}{2\sigma^2} \left\| \mathbf{\theta}\right\|_2^2 \right\},
\end{eqnarray*}
where $\mathbf{Y}$ is defined as $(Y_1, \dots, Y_m)^T \in \left\{ 0,1\right\}$ and $\sigma^2 $ is an additional parameter. We have
\begin{eqnarray*}
U( \mathbf{\theta}) = \mathbf{Y}^T \mathbf{X \theta} + \sum_{i=1}^m \text{log} (1 + e^{-\mathbf{\theta}^T \mathbf{X}_i}) + \frac{1}{2\sigma^2} \left\|\mathbf{\theta}\right\|_2^2, \,\,
\nabla U(\mathbf{\theta}) = \mathbf{X}^T \mathbf{Y} - \sum_{i=1}^m \frac{\mathbf{X}_i}{1 + e^{\mathbf{\theta}^T \mathbf{X}_i}} + \frac{1}{\sigma^2} \mathbf{\theta}.
\end{eqnarray*}
To demonstrate the performance of the proposed control variates approach in the above Bayesian logistic regression model, we take a simple dataset from \cite{mira2013zero}, which contains the measurements of four variables on $m=200$ Swiss banknotes. Prior distribution of the regression parameter $\mathbf{\theta} = \left( \theta_1, \theta_2, \theta_3, \theta_4 \right)$ assumed to be normal with the covariance matrix $\sigma^2 I_4$, where $\sigma^2 = 100$. To construct trajectories of length $n = 5 \times 10^3, $ we take  constant step sizes $\gamma_i = 0.1$ for the ULA scheme with $N = 10^3$ burn-in steps. As in the previous experiment we use the first order polynomials approximations to analytically compute the coefficients $\hat{a}_{p-l,\mathbf{k}},$ where $T = 10$ and $n_{trunc} = 50$. The target function is taken to be $f(\theta) = \sum_{i=1}^{i=d}\theta_i$.  In order to test our variance reduction algorithm, we generates \(100\) independent test trajectories.
In Figure ~\ref{fig:2} we compare our approach to variance reduction methods of \cite{mira2013zero} and \cite{brosse2018diffusion}.

\subsection{Binary Probit Regression}

Finally we consider a Bayesian probit regression model in its ordinary form, following \cite{mira2013zero}. More precisely, the log-likelihood of the model looks as follows
\begin{equation*}
    \mathbf{L}(\mathbf{Y}| \theta, \mathbf{X}) = \sum_{i=1}^{m} \left[ \mathbf{Y}_i \text{log} (\Phi(\theta^T \mathbf{X}_i)) + (1-\mathbf{Y}_i) \text{log}(\Phi(-\theta^T \mathbf{X}_i))\right],
\end{equation*}
where $\theta \in \mathbb{R}^d$ is a parameter of the model and $\Phi$ is a cumulative distribution function of the standard normal distribution $\mathcal{N}(0,1)$. Then, if the prior $\pi(\mathbf{\theta}) = \mathcal{N}(0, \sigma^2 I_d)$, then the posterior distribution expresses as
\begin{equation*}
    \pi(\theta | \mathbf{Y}, \mathbf{X})  \propto \text{exp} \left\{\sum_{i=1}^{m} \left[\mathbf{Y}_i \text{log} (\Phi(\theta^T \mathbf{X}_i)) + (1-\mathbf{Y}_i) \text{log}(\Phi(-\theta^T \mathbf{X}_i))\right]- \frac{1}{2\sigma^2} \left\| \mathbf{\theta}\right\|_2^2  \right\}
\end{equation*}
Consequently,
\begin{equation*}
    U(\theta) =  - \sum_{i=1}^{m} \left[\mathbf{Y}_i \text{log} (\Phi(\theta^T \mathbf{X}_i)) + (1-\mathbf{Y}_i) \text{log}(\Phi(-\theta^T \mathbf{X}_i))\right] + \frac{1}{2\sigma^2} \left\| \mathbf{\theta}\right\|_2^2
\end{equation*}

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.6\textwidth]{../pictures/Probit_banknotes_4d.png}
\caption{Boxplots of ergodic averages  from the variance reduced ULA algorithms for the Probit Regression. The compared estimators are the ordinary empirical average  (O), our modified estimator (CV-B), zero variance estimator (ZV) and the estimator with control variates using asymptotic variance minimization (CV).
 \label{fig:3}}
\end{center}
\end{figure}

\begin{equation*}
    \nabla U(\theta) = - \mathbf{X}^T s + \frac{1}{\sigma^2} \theta,
\end{equation*}
where the elements of the \(m\)-dimensional column vector $s$ are given by
\begin{equation*}
    s_i = \mathbf{Y}_i \Omega (\theta^T \mathbf{X}_i) - (1 - \mathbf{Y}_i) \Omega(- \theta^T \mathbf{X}_i), \quad i = 1, 2, \dots, m,
\end{equation*}
and $\Omega(\theta^T \mathbf{X}_i) : = \phi(\theta^T \mathbf{X}_i) / \Phi(\theta^T \mathbf{X}_i)$, where $\phi$ stands for the density of the standard normal distribution.
As in the previous examples, to show the performance  of the proposed variance reduced estimator for the Bayesian probit regression model, banknotes dataset is used and the regularization parameter is taken to be $\sigma^2=100$.
To generate trajectories of the length $n = 5 \times 10^3, $ we take a constant step size $\gamma_i = 0.1$ for the ULA scheme after $N = 10^3$ burn-in steps.  We use again the first order polynomials approximations to  compute the coefficients $\hat{a}_{p-l,\mathbf{k}}$ and fixed $T = 10$, $n_{trunc} = 50$ and $K=1$. The target function is now $f(\theta) = \sum_{i=1}^{i=d}\theta_i$.  In order to test our variance reduction algorithm, we generate \(100\) independent test trajectories. In Figure ~\ref{fig:3} we compare our approach to the variance reduction methods of \cite{mira2013zero} and \cite{brosse2018diffusion}.

\section{Proofs}
\label{sec:proofs}
\subsection{Proof of Theorem~\ref{th:mr}}
We start with introducing some notations.
For $m\in\mathbb N$, a smooth function
$h\colon\mathbb R^{d\times m}\to\mathbb R$
with arguments being denoted
$(y_1,\ldots,y_m)$, $y_i\in\mathbb R^d$, $i=1,\ldots,m$,
a multi-index $\mathbf k=(k_i)\in\mathbb N_0^d$,
and $j\in\{1,\ldots,m\}$,
we use the notation $\partial^{\mathbf k}_{y_j} h$ for the multiple derivative of $h$
with respect to the components of~$y_j$:
$$
\partial^{\mathbf k}_{y_j} h(y_1,\ldots,y_m)
:=\partial^{k_d}_{y_j^d}
\ldots
\partial^{k_1}_{y_j^1}
h(y_1,\ldots,z_m),
\quad y_j=(y_j^1,\ldots,y_j^d).
$$
In the particular case $m=1$ we can drop
the subscript $y_1$ in that notation.
For $l\le p$, we have the representation
\[
X_{p}=G_{p,l}(X_{l-1},\sqrt{\gamma_{l}}Z_{l},\ldots,\sqrt{\gamma_{p}}Z_{p}),
\]
where the function \(G_{p,l}:\) \(\mathbb{R}^{d\times(p-l+2)}\to \mathbb{R}^{d}\) is defined as
\begin{equation}
\label{eq:definition-G-p-l}
G_{p,l}(x,y_l,\ldots,y_p):=\Phi_{p}(\cdot,y_{p})\circ\Phi_{p-1}(\cdot,y_{p-1})\circ\ldots\circ\Phi_{l}(x,y_{l})
\end{equation}
with
\[
\Phi_{j}(x,y)=x-\gamma_{j}\mu(x)+y,
\quad x,y\in\mathbb R^d.
\]
As a consequence,
for a function $f\colon\mathbb R^d\to\mathbb R$
as in Section~\ref{sec:setup}, we have
$$
f\left(X_{p}\right) =f\circ G_{p,l}(X_{l-1},\sqrt{\gamma_{l}}Z_{l},\ldots,\sqrt{\gamma_{p}}Z_{p}).
$$
In what follows, for $\mathbf{k}\in\mathbb N_0^d$,
we use the shorthand notation
\begin{equation}
\label{eq:definition-differential-f-p}
\partial_{y_l}^{\mathbf k} f\left(X_{p}\right)
:=\partial_{y_l}^{\mathbf k} [f\circ G_{p,l}](X_{l-1},\sqrt{\gamma_{l}}Z_{l},\ldots,\sqrt{\gamma_{p}}Z_{p})
\end{equation}
whenever the function $f\circ G_{p,l}$ is smooth enough
(that is, $f$ and $\mu$ need to be smooth enough).
Finally, for a multi-index $\mathbf{k}=(k_i)\in\mathbb N_0^d$, we use the notation
\(\mathbf{k} ! :=k_1!\cdot\ldots\cdot k_d !\)

\begin{lem}\label{eq:a_repr}
Fix $l\le p$ and some \(\mathbf{k},\mathbf{k}'\in \mathbb{N}_0^d\) with \(\mathbf{k}'\le \mathbf{k}\) componentwise.  Then the following representation holds
$$
a_{p,l,\mathbf{k}}(X_{l-1})=\left(\gamma_{l}^{|\mathbf{k}'|}\frac{(\mathbf{k}-\mathbf{k}') !}{\mathbf{k}!}\right)^{1/2}
\mathsf E\left[\left.
\partial_{y_l}^{\mathbf{k}'}f(X_p)\mathbf{H}_{\mathbf{k}-\mathbf{k}'}(Z_l)\right|X_{l-1}\right].
$$
\end{lem}

\begin{proof}
Let $\boldsymbol{\varphi}(z)=\frac{1}{(2\pi)^{d/2}} \exp\{-|z|^2/2\}$,
$z\in\mathbb R^d$, denote the density of a $d$-dimensional
standard Gaussian random vector.
We first remark that, for the normalized Hermite polynomial $\mathbf H_{\mathbf k}$ on $\mathbb R^d$,
$\mathbf k\in\mathbb N_0^d$, it holds
$$
\mathbf{H}_{\mathbf{k}}(z)\boldsymbol{\varphi}(z)
=\frac{(-1)^{|\mathbf{k}|}}{\sqrt{\mathbf{k} !}} \partial^{\mathbf{k}} \boldsymbol{\varphi}(z).
$$
This enables to use the integration by parts in vector form as follows
(below $\prod_{j=l+1}^p:=1$ whenever $l=p$)
\begin{align*}
&a_{p,l,\mathbf k}(x)
 =
\int_{\mathbb R^d}\ldots\int_{\mathbb R^d}
f\circ G_{p,l}(x,\sqrt{\gamma_{l}}z_{l},\ldots,\sqrt{\gamma_{p}}z_{p})
\mathbf{H}_{\mathbf{k}}(z_{l})\boldsymbol{\varphi}(z_l)
\prod_{j=l+1}^p\boldsymbol{\varphi}(z_j)\, dz_{l}\ldots dz_{p}
\\
& =
\frac{1}{\sqrt{\mathbf k!}}
\int_{\mathbb R^d}\ldots\int_{\mathbb R^d}
f\circ G_{p,l}(x,\sqrt{\gamma_{l}}z_{l},\ldots,\sqrt{\gamma_{p}}z_{p})
(-1)^{|\mathbf{k}|}\partial^{\mathbf{k}} \boldsymbol{\varphi}(z_l)
\prod_{j=l+1}^p\boldsymbol{\varphi}(z_j)\, dz_{l}\ldots dz_{p}
\\
& =
\gamma_l^{|\mathbf k'|/2}\frac{1}{\sqrt{\mathbf k!}}
\int_{\mathbb R^d}\ldots\int_{\mathbb R^d}
\partial_{y_l}^{\mathbf k'}[f\circ G_{p,l}](x,\sqrt{\gamma_{l}}z_{l},\ldots,\sqrt{\gamma_{p}}z_{p})
\\
&\hspace{1em}\times
(-1)^{|\mathbf{k}-\mathbf k'|}\partial^{\mathbf{k}-\mathbf k'} \boldsymbol{\varphi}(z_l)
\prod_{j=l+1}^p\boldsymbol{\varphi}(z_j)\, dz_{l}\ldots dz_{p}
\\
%& =
%\gamma_l^{|\mathbf k'|/2}\frac{\sqrt{(\mathbf k-\mathbf k')!}}{\sqrt{\mathbf k!}}
%\int_{\mathbb R^d}\ldots\int_{\mathbb R^d}
%\partial_{y_l}^{\mathbf k'}[f\circ G_{p,l}](x,\sqrt{\gamma_{l}}z_{l},\ldots,\sqrt{\gamma_{p}}z_{p})
%\\
&\hspace{1em}\times
\mathbf{H}_{\mathbf{k}-\mathbf k'}(z_{l})\boldsymbol{\varphi}(z_l)
\prod_{j=l+1}^p\boldsymbol{\varphi}(z_j)\, dz_{l}\ldots dz_{p}
\\
 & =\gamma_{l}^{|\mathbf{k}'|/2}\frac{\sqrt{(\mathbf{k}-\mathbf{k}')!}}{\sqrt{\mathbf{k}!}}\mathsf{E}\left[\partial_{y_{l}}^{\mathbf{k}'}[f\circ G_{p,l}](x,\sqrt{\gamma_{l}}Z_{l},\ldots,\sqrt{\gamma_{p}}Z_{p})\mathbf{H}_{\mathbf{k}-\mathbf{k}'}(Z_{l})\right].
\end{align*}
The last expression yields the result.
\end{proof}

For multi-indices $\mathbf k,\mathbf k'\in\mathbb N_0^d$
with $\mathbf k'\le\mathbf k$ componentwise
and $\mathbf k'\ne\mathbf k$,
we get applying first Lemma~\ref{eq:a_repr}
\begin{equation*}
\begin{split}
\bar a_{l,\mathbf{k}}(X_{l-1})
&=\left(\gamma_{l}^{|\mathbf{k}'|}\frac{(\mathbf{k}-\mathbf{k}')!}{\mathbf{k}!}\right)^{1/2}
\,\sum_{p=l}^{N+n}\omega_{p,n}^N
\mathsf E\left[\left.
\partial_{y_l}^{\mathbf{k}'}f(X_p)\mathbf{H}_{\mathbf{k}-\mathbf{k}'}(Z_l)\right|X_{l-1}\right]\\
&=\left(\gamma_{l}^{|\mathbf{k}'|}\frac{(\mathbf{k}-\mathbf{k}')!}{\mathbf{k}!}\right)^{1/2}
\\
&\times \sum_{p=l}^{N+n}\omega_{p,n}^N
\mathsf E\left[\left.
\left(
\partial_{y_l}^{\mathbf{k}'}f(X_p)-\mathsf E\left[\partial_{y_l}^{\mathbf{k}'}f(X_p)|X_{l-1}\right]
\right)
\mathbf{H}_{\mathbf{k}-\mathbf{k}'}(Z_l)\right|X_{l-1}\right].
\end{split}
\end{equation*}

Assume that $\mu$ and $f$ are $K\times d$ times continuously differentiable.
Then, given $\mathbf k\in\mathbb N_0^d$,
by taking $\mathbf k'= \mathbf k'(\mathbf k)
=(K1_{\{k_1>K\}},\ldots,K1_{\{k_d>K\}})$,
for each $l\in\{N+1,\ldots,N+n\}$, we get
\begin{align}
\label{eq:sum_abar}
\sum_{\mathbf{k}\colon\|\mathbf{k}\|\geq K+1}\left[\bar{a}_{l,\mathbf{k}}^{2}(X_{l-1})|\mathcal{G}_{N}\right] & =\sum_{\mathbf{k}\colon\|\mathbf{k}\|\geq K+1}\left(\gamma_{l}^{|\mathbf{k}'|}\frac{(\mathbf{k}-\mathbf{k}')!}{\mathbf{k}!}\right)Q(\mathbf{k}',\mathbf{k}-\mathbf{k}')\\
\nonumber
 & =\sum_{I\subseteq\{1,\ldots,d\},\, I\neq \emptyset}\gamma_{l}^{|I|K}\sum_{\mathbf{m}_{I}\in\mathbb{N}_{I}^{d}}\frac{\mathbf{m}_{I}!}{\left(\mathbf{m}_{I}+\mathbf{K}_{I}\right)!}
 \\
 \nonumber
 & \hspace{1em}\times\sum_{\mathbf{m}_{I^c}\in \mathbb{N}^d_{0,I^c},\,\|\mathbf{m}_{I^c}\|\leq K}Q(\mathbf{\mathbf{K}}_{I},\mathbf{m}_{I}+\mathbf{m}_{I^c}),
\end{align}
where for any two multi-indices \(\mathbf{r},\) \(\mathbf{q}\) from \(\mathbb{N}_0^d\)
\begin{multline*}
Q(\mathbf{r},\mathbf{q})
\\
=\mathsf{E}\left\{
\left(\mathsf{E}\left[\sum_{p=l}^{N+n}\omega_{p,n}^{N}\left(\partial_{y_{l}}^{\mathbf r}f\left(X_{p}\right)-\mathsf{E}\left[\partial_{y_{l}}^{\mathbf r}f\left(X_{p}\right)|X_{l-1}\right]\right)\mathbf{H}_{\mathbf{q}}(Z_{l})\Big|X_{l-1}\right]\right)^{2}
\right\}.
\end{multline*}
In \eqref{eq:sum_abar} the first sum runs over all nonempty subsets $I$ of the set $\{1,\ldots,d\}.$
For any subset $I,$ $\mathbb{N}_{I}^{d}$ stands for a set
of multi-indices $\mathbf{m}_{I}$ with elements $m_{i}=0,$ $i\not\in I,$
and $m_{i}\in\mathbb{N},$  $i\in I.$ Moreover, \(I^c=\{1,\ldots,d\}\setminus I\) and \(\mathbb{N}^d_{0,I^c}\) stands for a set
of multi-indices $\mathbf{m}_{I^c}$ with elements $m_{i}=0,$ $i\in I,$
and $m_{i}\in\mathbb{N}_0,$  $i\not\in I.$ Finally, the multi-index \(\mathbf{K}_I\) is defined as $\mathbf{\mathbf{K}}_{I}=(K1_{\{1\in I\}},\ldots,K1_{\{d\in I\}}).$
Applying the estimate
\begin{eqnarray*}
\frac{\mathbf{m}_{I}!}{\left(\mathbf{m}_{I}+\mathbf{K}_{I}\right)!}\leq (1/2)^{|I| K},
\end{eqnarray*}
we get
\begin{eqnarray*}
\sum_{\mathbf{k}\colon\|\mathbf{k}\|\geq K+1}\left[\bar{a}_{l,\mathbf{k}}^{2}(X_{l-1})|\mathcal{G}_{N}\right]&\leq &
\sum_{I\subseteq\{1,\ldots,d\},\, I\neq \emptyset} (\gamma_{l}/2)^{|I|K}
\\
&& \times\sum_{\mathbf{m}_{I}\in\mathbb{N}_{I}^{d}} \sum_{\mathbf{m}_{I^c}\in \mathbb{N}^d_{0,I^c},\,\|\mathbf{m}_{I^c}\|\leq K} Q(\mathbf{\mathbf{K}}_{I},\mathbf{m}_{I}+\mathbf{m}_{I^c})
\\
&\leq &
\sum_{I\subseteq\{1,\ldots,d\},\, I\neq \emptyset} (\gamma_{l}/2)^{|I|K} \sum_{\mathbf{m}\in\mathbb{N}_0^{d}} Q(\mathbf{\mathbf{K}}_{I},\mathbf{m}).
\end{eqnarray*}
Now using the consequence
$\sum_{\mathbf{m}\in \mathbb{N}^d_0} \langle\xi,\mathbf{H}_\mathbf{m}(Z_l)\rangle^2\le\langle\xi,\xi\rangle$
of Parseval's identity
(the latter is used conditionally on $X_{l-1}$
%equivalently, on $\mathcal G_{l-1}$,
which is possible because the system
$\{\mathbf{H}_\mathbf{m}(Z_l)\}_{m\in \mathbb{N}^d_0}$
is orthonormal conditionally on $X_{l-1}$),
%and the tower property $\mathsf E[\mathsf E(\cdot|\mathcal G_{l-1})|\mathcal G_N]=\mathsf E[\cdot|\mathcal G_N]$
we derive
\begin{eqnarray*}
\sum_{\mathbf{k}\colon\|\mathbf{k}\|\geq K+1}\left[\bar{a}_{l,\mathbf{k}}^{2}(X_{l-1})|\mathcal{G}_{N}\right]
&\leq & \sum_{I\subseteq\{1,\ldots,d\},\, I\neq \emptyset}
\left(\frac{\gamma_{l}}{2}\right)^{|I|K}
\\
&& \times\mathsf E\left[
\mathsf{Var}\left(\sum_{p=l}^{N+n}\omega_{p,n}^{N}\partial_{y_{l}}^{\mathbf{K}_I}f\left(X_{p}\right)
\Big|X_{l-1}\right)
\Bigg|\mathcal G_N\right]
\\
&=& \frac{1}{\Gamma^2_{N+2,N+n+1}}\sum_{I\subseteq\{1,\ldots,d\},\, I\neq \emptyset}
\left(\frac{\gamma_{l}}{2}\right)^{|I|K}
R_{l,n,N}^{I,K}
\end{eqnarray*}
with
\begin{eqnarray*}
R_{l,n,N}^{I,K}=\mathsf E\left[
\mathsf{Var}\left(\sum_{p=l}^{N+n}\gamma_{p+1}\partial_{y_{l}}^{\mathbf{K}_I}f\left(X_{p}\right)
\Big|X_{l-1}\right)\right].
\end{eqnarray*}
As a result
\begin{eqnarray*}
\mathsf{Var}\left[\left.\pi_{K,n}^{N}(f)\right|\mathcal G_N\right]\leq \frac{1}{\Gamma^2_{N+2,N+n+1}}\sum_{l=N+1}^{N+n}\sum_{I\subseteq\{1,\ldots,d\},\, I\neq \emptyset}
\left(\frac{\gamma_{l}}{2}\right)^{|I|K}
R_{l,n,N}^{I,K}.
\end{eqnarray*}
Next we show that under the conditions  of Theorem~\ref{th:mr}, the quantity \(R_{l,n,N}^{I,K}\) is uniformly bounded in \(l,n,N,I.\) For the sake of simplicity we present the proof only in one-dimensional case. First we need to prove several auxiliary results.
\begin{lem}\label{lem:06062018a1}
Let $(x_p)_{p\in\mathbb N_0}$
and $(\epsilon_p)_{p\in\mathbb N}$
be sequences of nonnegative real numbers
satisfying $x_0=\ol C_0$ and
\begin{align}
0&\le x_p\le\alpha_p x_{p-1}+\gamma_p \epsilon_p,\quad p\in\mathbb N,
\label{eq:06062018a1}\\
0&\le\epsilon_p\le\ol C_1\prod_{k=1}^p \alpha_k^2,\quad p\in\mathbb N,
\label{eq:06062018a2}
\end{align}
where $\alpha_p,\gamma_p\in(0,1)$, $p\in\mathbb N$,
and $\ol C_0,\ol C_1$ are some nonnegative constants. Assume
\begin{equation}\label{eq:06062018a3}
\sum_{r=1}^\infty\gamma_r \prod_{k=1}^r \alpha_k\le\ol C_2
\end{equation}
for some constant $\ol C_2$. Then
$$
x_p\le(\ol C_0+\ol C_1\ol C_2)\prod_{k=1}^p \alpha_k,\quad p\in\mathbb N.
$$
\end{lem}

\begin{proof}
Applying~\eqref{eq:06062018a1} recursively, we get
$$
x_p\le\ol C_0\prod_{k=1}^p \alpha_k
+\sum_{r=1}^p \gamma_r\epsilon_r
\prod_{k=r+1}^p \alpha_k,
$$
where we use the convention $\prod_{k=p+1}^p:=1$.
Substituting estimate~\eqref{eq:06062018a2}
into the right-hand side, we obtain
$$
x_p\le\left(\ol C_0+\ol C_1
\sum_{r=1}^p \gamma_r \prod_{k=1}^r \alpha_k
\right)
\prod_{k=1}^p \alpha_k,
$$
which, together with~\eqref{eq:06062018a3}, completes the proof.
\end{proof}

In what follows, we use the notation
\begin{equation}
\label{eq:definition-alpha}
\alpha_k=1-\gamma_k \mu'(X_{k-1}),\quad k\in\mathbb N.
\end{equation}

\begin{remark}\label{rem:10062018a1}

Below this will allow us to apply
Lemma~\ref{lem:06062018a1}
to bound double indexed sequences
$(x_{j,p})_{j\ge1,p\ge j}$
satisfying
$$
0\le x_{j,p}\le\alpha_p x_{j,p-1}+\gamma_p \epsilon_{j,p},\quad p\ge j+1,
$$
with suitable $(\epsilon_{j,p})_{j\ge1,p\ge j+1}$.
\end{remark}
The assumption (H1) implies that  $|\mu'(x)|\leq B_\mu$ for some constant $B_\mu>0$ and all $x\in \mathbb{R}^d$. Without loss of generality we suppose that $\gamma_N B_\mu<1.$
\begin{lem}\label{lem:06062018a2}
Under assumptions of Theorem~\ref{th:mr},
for all natural $r\le K$ and $l\le p$, there exist constants $C_r$ (not depending on $l$ and $p$) such that
\begin{equation}
\label{eq:08062018b2}
\left|\partial_{y_l}^r X_p \right| \le C_r\prod_{k=l+1}^p (1-\gamma_k\mu'(X_{k-1}))\quad \text{a.s.} 
\end{equation}
where $\partial_{y_l}^r X_p$ is defined in \eqref{eq:definition-differential-f-p}. Moreover, we can choose $C_1=1$.
\end{lem}
\begin{lem}\label{lem:06062018a3}
Under assumptions of Theorem~\ref{th:mr},
for all natural $r\le K$, $j\ge l$ and $p>j$, we have
\begin{equation}\label{eq:08062018b3}
\left|\partial_{y_j} \partial_{y_l}^r X_p\right|
\le c_r\prod_{k=l+1}^p (1-\gamma_k\mu'(X_{k-1})), \quad \text{a.s.}
\end{equation}
with some constants $c_r$
not depending on $j$, $l$ and $p$,
while, for $p\le j$, it holds
$\partial_{y_{j}}\partial_{y_{l}}^{r}X_{p}=0$.
\end{lem}

\begin{proof}
The last statement is straightforward.
We fix natural numbers $j\ge l$ and prove~\eqref{eq:08062018b3}
for all $p>j$ by induction in~$r$.
First, for $p>j$, we write
$$
\partial_{y_{l}}X_{p}
=\left[1-\gamma_{p}\mu'(X_{p-1})\right]\partial_{y_{l}}X_{p-1}
$$
and differentiate this identity with respect to~$y_j$
$$
\partial_{y_{j}}\partial_{y_{l}}X_{p}
=\left[1-\gamma_{p}\mu'(X_{p-1})\right]\partial_{y_{j}}\partial_{y_{l}}X_{p-1}-\gamma_{p}\mu''(X_{p-1})\partial_{y_{j}}X_{p-1}\partial_{y_{l}}X_{p-1}.
$$
By Lemma~\ref{lem:06062018a2}, we have
\begin{align*}
|\partial_{y_{j}}\partial_{y_{l}}X_{p}|
&\le\alpha_p|\partial_{y_{j}}\partial_{y_{l}}X_{p-1}|
+\gamma_{p}
B_\mu
\prod_{k=l+1}^{p-1}\alpha_k
\prod_{k=j+1}^{p-1}\alpha_k\\
&\le\alpha_p|\partial_{y_{j}}\partial_{y_{l}}X_{p-1}|
+\gamma_{p}
\const
\prod_{k=l+1}^{j}\alpha_k
\prod_{k=j+1}^{p}\alpha_k^2,
\quad p\ge j+1,
\end{align*}
with a suitable constant.
By Lemma~\ref{lem:06062018a1}
together with Remark~\ref{rem:10062018a1} applied
to bound $|\partial_{y_{j}}\partial_{y_{l}}X_{p}|$
for $p\ge j+1$
(notice that $\partial_{y_j}\partial_{y_l}X_j=0$, that is,
$\ol C_0$ in Lemma~\ref{lem:06062018a1} is zero,
while $\ol C_1$ in Lemma~\ref{lem:06062018a1}
has the form $\const\prod_{k=l+1}^j \alpha_k$),
we obtain~\eqref{eq:08062018b3} for $r=1$.
The induction hypothesis is now that the inequality
\begin{equation}\label{eq:induc-derive}
\left|\partial_{y_{j}}\partial_{y_{l}}^{k}X_{p}\right|
\leq c_{k}\prod_{s=l+1}^{p}\alpha_{s}
\end{equation}
holds for all natural $k<r\;(\le K)$ and $p>j$.
We need to show~\eqref{eq:induc-derive} for $k=r$.
Fa\`a di Bruno's formula implies for $2\le r\le K$ and $p>l$
\begin{align}
\partial_{y_{l}}^{r}X_{p}
&=\left[1-\gamma_{p}\mu'(X_{p-1})\right]\partial_{y_{l}}^{r}X_{p-1}
\label{eq:08062018a1}\\
&\hspace{1em}-\gamma_{p}\sum\frac{r!}{m_{1}!\ldots m_{r-1}!\,}\mu^{(m_{1}+\ldots+m_{r-1})}(X_{p-1})\prod_{k=1}^{r-1}\left(\frac{\partial_{y_{l}}^{k}X_{p-1}}{k!}\right)^{m_{k}},
\notag
\end{align}
where the sum is taken over all $(r-1)$-tuples of nonnegative integers
$(m_{1},\ldots,m_{r-1})$ satisfying the constraint
\begin{equation}\label{eq:09062018a1}
1\cdot m_{1}+2\cdot m_{2}+\ldots+(r-1)\cdot m_{r-1}=r.
\end{equation}
Notice that we work with $(r-1)$-tuples
rather than with $r$-tuples
because the term containing
$\partial_{y_l}^r X_{p-1}$
on the right-hand side of~\eqref{eq:08062018a1}
is listed separately.
For $p>j$, we then have
\begin{align}
&\partial_{y_{j}}\partial_{y_{l}}^{r}X_{p}
=\left[1-\gamma_{p}\mu'(X_{p-1})\right]\partial_{y_{j}}\partial_{y_{l}}^{r}X_{p-1}-\gamma_{p}\mu''(X_{p-1})\partial_{y_{l}}^{r}X_{p-1}\partial_{y_{j}}X_{p-1}
\label{eq:09062018a2}\\
&\hspace{1em}-\gamma_{p}\sum\frac{r!}{m_{1}!\ldots m_{r-1}!\,}\mu^{(m_{1}+\ldots+m_{r-1}+1)}(X_{p-1})\partial_{y_{j}}X_{p-1}\prod_{k=1}^{r-1}\left(\frac{\partial_{y_{l}}^{k}X_{p-1}}{k!}\right)^{m_{k}}
\notag\\
&\hspace{1em}-\gamma_{p}\sum\frac{r!}{m_{1}!\ldots m_{r-1}!\,}\mu^{(m_{1}+\ldots+m_{r-1})}(X_{p-1})\partial_{y_{j}}\left[\prod_{k=1}^{r-1}\left(\frac{\partial_{y_{l}}^{k}X_{p-1}}{k!}\right)^{m_{k}}\right]
\notag\\
&\hspace{1em} =\left[1-\gamma_{p}\mu'(X_{p-1})\right]\partial_{y_{j}}\partial_{y_{l}}^{r}X_{p-1}+\gamma_{p}\epsilon_{l,j,p},
\notag
\end{align}
where the last equality defines the quantity $\epsilon_{l,j,p}$.
Furthermore,
\begin{eqnarray*}
\partial_{y_{j}}\left[\prod_{k=1}^{r-1}\left(\frac{\partial_{y_{l}}^{k}X_{p-1}}{k!}\right)^{m_{k}}\right]&=&\sum_{s=1}^{r-1}\frac{m_{s}}{s!}\left(\frac{\partial_{y_{l}}^{s}X_{p-1}}{s!}\right)^{m_{s}-1}\partial_{y_{j}}\partial_{y_{l}}^{s}X_{p-1}
\\
&& \times \prod_{k\le r-1,\,k\neq s}\left(\frac{\partial_{y_{l}}^{k}X_{p-1}}{k!}\right)^{m_{k}}.
\end{eqnarray*}
Using Lemma~\ref{lem:06062018a2},
induction hypothesis~\eqref{eq:induc-derive}
and the fact that $m_{1}+\ldots+m_{r-1}\ge2$
for $(r-1)$-tuples of nonnegative integers
satisfying~\eqref{eq:09062018a1},
we can bound $|\epsilon_{l,j,p}|$ as follows
\begin{align*}
&\left|\epsilon_{l,j,p}\right|  \leq  B_{\mu}C_{r}\prod_{k=l+1}^{p-1}\alpha_{k}\prod_{k=j+1}^{p-1}\alpha_{k}+B_{\mu}\sum\frac{r!}{m_{1}!\ldots m_{r-1}!\,}\left[\prod_{k=j+1}^{p-1}\alpha_{k}\right]
\\
&\times \prod_{s=1}^{r-1}\left(\frac{C_{s}\prod_{k=l+1}^{p-1}\alpha_{k}}{s!}\right)^{m_{s}}\\
&+B_{\mu}\sum\frac{r!}{m_{1}!\ldots m_{r-1}!\,}\sum_{t=1}^{r-1}\frac{m_t}{t!}\left(\frac{C_{t}\prod_{k=l+1}^{p-1}\alpha_{k}}{t!}\right)^{m_{t}-1}c_t\left[\prod_{k=l+1}^{p-1}\alpha_{k}\right]
\\
&\times \prod_{s\le r-1,\,s\neq t}\left(\frac{C_{s}\prod_{k=l+1}^{p-1}\alpha_{k}}{s!}\right)^{m_{s}}\leq \const\prod_{k=l+1}^{j}\alpha_{k}\prod_{k=j+1}^{p}\alpha_{k}^{2}
\end{align*}
with some constant ``$\const$'' depending on
$B_\mu,r,C_1,\ldots,C_r,c_1,\ldots,c_{r-1}$.
Thus, \eqref{eq:09062018a2} now implies
$$
|\partial_{y_{j}}\partial_{y_{l}}^{r}X_{p}|
\le\alpha_p|\partial_{y_{j}}\partial_{y_{l}}^{r}X_{p-1}|
+\gamma_p \const\prod_{k=l+1}^{j}\alpha_{k}\prod_{k=j+1}^{p}\alpha_{k}^{2},
\quad p\ge j+1.
$$
We can again apply Lemma~\ref{lem:06062018a1}
and Remark~\ref{rem:10062018a1}
to bound $|\partial_{y_{j}}\partial_{y_{l}}^r X_{p}|$
for $p\ge j+1$
(notice that $\partial_{y_j}\partial_{y_l}^r X_j=0$, that is,
$\ol C_0$ in Lemma~\ref{lem:06062018a1} is zero,
while $\ol C_1$ in Lemma~\ref{lem:06062018a1}
has the form $\const\prod_{k=l+1}^j \alpha_k$),
and we obtain~\eqref{eq:induc-derive} for $k=r$.
This concludes the proof.
\end{proof}

\begin{lem}
\label{lem:var_poincare}
Under assumptions of Theorem~\ref{th:mr},
for all natural $l\le q$, it holds
$$
\mathsf{Var}\left[\sum_{p=l}^{q}\gamma_{p+1}\partial_{y_{l}}^{K}f\left(X_{p}\right)\Big|X_{l-1}\right]\le B_K\quad\text{a.s.},
$$
where $B_K$ is a deterministic bound that does not depend on $l$ and~$q$.
\end{lem}

\begin{proof}
The expression
$\sum_{p=l}^q \gamma_{p+1} \partial_{y_l}^K f(X_p)$
can be viewed as a deterministic function of
$X_{l-1},Z_l,Z_{l+1},\ldots,Z_q$
$$
\sum_{p=l}^q \gamma_{p+1} \partial_{y_l}^K f(X_p)
=F(X_{l-1},Z_l,Z_{l+1},\ldots,Z_q).
$$
By the (conditional) Gaussian Poincar\'e inequality,
we have
$$
\mathsf{Var}\left[\sum_{p=l}^{q}\gamma_{p+1}\partial_{y_{l}}^{K}f\left(X_{p}\right)\Big|X_{l-1}\right]
\le\mathsf E\left[\left.
\|\nabla_Z F(X_{l-1},Z_l,Z_{l+1},\ldots,Z_q)\|^2
\right|X_{l-1}\right],
$$
where $\nabla_Z F=(\partial_{Z_l} F,\ldots,\partial_{Z_q} F)$,
and $\|\cdot\|$ denotes the Euclidean norm.
Notice that \(\partial_{Z_j} F=\sqrt{\gamma_j}\,\partial_{y_j} F\) and
hence
$$
\mathsf{Var}\left[\sum_{p=l}^{q}\gamma_{p+1}\partial_{y_{l}}^{K}f\left(X_{p}\right)\Big|X_{l-1}\right]\leq\sum_{j=l}^{q}\gamma_{j}\mathsf{E}\left[\left(\sum_{p=l}^{q}\gamma_{p+1}\partial_{y_{j}}\partial_{y_{l}}^{K}f\left(X_{p}\right)\right)^{2}\Big|X_{l-1}\right].
$$
It is straightforward to check that
$\partial_{y_j}\partial_{y_l}^K f(X_p)=0$
whenever $p<j$. Therefore, we get
\begin{equation}\label{eq:10062018a2}
\mathsf{Var}\left[\sum_{p=l}^{q}\gamma_{p+1}\partial_{y_{l}}^{K}f\left(X_{p}\right)\Big|X_{l-1}\right]\leq\sum_{j=l}^{q}\gamma_{j}\mathsf{E}\left[\left(\sum_{p=j}^{q}\gamma_{p+1}\partial_{y_{j}}\partial_{y_{l}}^{K}f\left(X_{p}\right)\right)^{2}\Big|X_{l-1}\right].
\end{equation}
Now fix $p$ and $j$, $p\ge j$, in $\{l,\ldots,q\}$.
By Fa\`a di Bruno's formula
\[
\partial_{y_{l}}^{K}f\left(X_{p}\right)=\sum\frac{K!}{m_{1}!\ldots m_{K}!}f^{(m_{1}+\ldots+m_{K})}(X_{p})\prod_{k=1}^{K}\left(\frac{\partial_{y_{l}}^{k}X_{p}}{k!}\right)^{m_{k}},
\]
where the sum is taken over all $K$-tuples of nonnegative integers
$(m_1,\ldots,m_K)$ satisfying
$$
1\cdot m_{1}+2\cdot m_{2}+\ldots+K\cdot m_{K}=K.
$$
Then
\begin{eqnarray*}
\partial_{y_{j}}\partial_{y_{l}}^{K}f\left(X_{p}\right)
&=&\sum\frac{K!}{m_{1}!\ldots m_{K}!}f^{(m_{1}+\ldots+m_{K}+1)}(X_{p})\left[\partial_{y_{j}}X_{p}\right]\prod_{k=1}^{K}\left(\frac{\partial_{y_{l}}^{k}X_{p}}{k!}\right)^{m_{k}} \\
&&+\sum\frac{K!}{m_{1}!\ldots m_{K}!}f^{(m_{1}+\ldots+m_{K})}(X_{p})
\sum_{s=1}^{K}\frac{m_{s}}{s!}
\left(\frac{\partial_{y_{l}}^{s}X_{p}}{s!}\right)^{m_{s}-1}
\\
&& \times\left[\partial_{y_{j}}\partial_{y_{l}}^{s}X_{p}\right]
\prod_{k\le K,\,k\neq s}\left(\frac{\partial_{y_{l}}^{k}X_{p}}{k!}\right)^{m_{k}}.
\end{eqnarray*}
Using the bounds of
Lemmas \ref{lem:06062018a2} and~\ref{lem:06062018a3},
we obtain
\begin{eqnarray}
\label{eq:part-part}
\left|\partial_{y_{j}}\partial_{y_{l}}^{K}f\left(X_{p}\right)\right|
\leq A_{K}\prod_{k=l+1}^{p}\alpha_{k}
\end{eqnarray}
with a suitable constant $A_{K}$.
Substituting this in~\eqref{eq:10062018a2},
we proceed as follows
\begin{align*}
\mathsf{Var}\left[\sum_{p=l}^{q}\gamma_{p+1}\partial_{y_{l}}^{K}f\left(X_{p}\right)\Big|X_{l-1}\right]
&\le
A_{K}^{2}\sum_{j=l}^{q}\gamma_{j}
\mathsf{E}\left(\sum_{p=j}^{q}\gamma_{p+1}\prod_{k=l+1}^{p}\alpha_{k}\right)^{2}
\\
&\le
\frac{A_{K}^{2}}{(1-\gamma_1 B_\mu)^2}
\mathsf{E}\sum_{j=l}^{q} \gamma_{j}
\left(
\sum_{p=j+1}^{q+1} \gamma_{p} \prod_{k=l+1}^{p} \alpha_{k}
\right)^{2}
\\
%&=
%\frac{A_{K}^{2}}{(1-\gamma_1 b_\mu)^2}
%\sum_{j=l}^{q} \gamma_{j}
%\prod_{k=l+1}^{j} \alpha_{k}^2
%\left(
%\sum_{p=j+1}^{q+1} \gamma_{p} \prod_{k=j+1}^{p} \alpha_{k}
%\right)^{2}
%\\
&\le
\frac{A_{K}^{2}}{(1-\gamma_1 B_\mu)^3}
\mathsf{E}\sum_{j=l}^{q} \gamma_{j}
\prod_{k=l}^{j} \alpha_{k}
\left(
\sum_{p=j+1}^{q+1} \gamma_{p} \prod_{k=j+1}^{p} \alpha_{k}
\right)^{2}
\end{align*}

Now, from Hoelder inequality, we obtain (with $\|X\|_p = (\mathsf{E}X^p)^{\frac{1}{p}}$)
\begin{align*}
\mathsf{E}\left[\sum_{j=l}^{q}\gamma_{j}\prod_{k=l}^{j}\alpha_{k}\left(\sum_{p=j+1}^{q+1}\gamma_{p}\prod_{k=j+1}^{p}\alpha_{k}\right)^{2}\right] & \leq\sum_{j=l}^{q}\gamma_{j}\left\Vert \prod_{k=l}^{j}\alpha_{k}\right\Vert _{2}\left\Vert \sum_{p=j+1}^{q+1}\gamma_{p}\prod_{k=j+1}^{p}\alpha_{k}\right\Vert _{4}^{2}\\
 & \leq\sum_{j=l}^{q}\gamma_{j}\left\Vert \prod_{k=l}^{j}\alpha_{k}\right\Vert _{2}\left(\sum_{p=j+1}^{q+1}\gamma_{p}\left\Vert \prod_{k=j+1}^{p}\alpha_{k}\right\Vert _{4}\right)^{2}.
\end{align*}
Now using the fact that $\prod\limits_{k=j+1}^{p}\alpha_{k} \leq \exp\left(-\sum\limits_{k=j+1}^{p}\gamma_{k}\mu'(X_{k-1})\right),$
we get 
\[
\mathsf{E}\left[\sum_{j=l}^{q}\gamma_{j}\prod_{k=l}^{j}\alpha_{k}\left(\sum_{p=j+1}^{q+1}\gamma_{p}\prod_{k=j+1}^{p}\alpha_{k}\right)^{2}\right]\leq\sum_{j=l}^{q}\gamma_{j} \zeta^{1/2}_{l,j}(2) \left(\sum_{p=j+1}^{q+1}\gamma_{p} \zeta^{1/4}_{j+1,p}(4)\right)^{2},
\]
where we denote
\[
\zeta_{l,j}(s)=\mathsf{E}\left[e^{-s\sum_{k=l}^{j}\gamma_{k}\mu'(X_{k-1})}\right],\quad s>0.
\]
From Theorem~\ref{exp_mom_ula} in Appendix~\ref{sec:appendix}, it follows that
\[
\mathsf{E}\left[e^{-s\left(\sum_{k=l}^{j}\gamma_{k}\mu'(X_{k-1})-\Gamma_{l,j}\int\mu'(x)\pi(x)\,dx\right)}\right]\leq e^{s\kappa\sum_{k=l}^{j}\gamma_{k}^{2}}
\]
with constant $\kappa = ...$. We used the fact that $\mu^{\prime}(x)$ is bounded. Denote $\sum_{k=1}^{\infty}\gamma_{k}^{2} = \Gamma^{\star}$, then we have 
\[
\mathsf{E}\left[e^{-s\sum_{k=l}^{j}\gamma_{k}\mu'(X_{k-1})}\right]\leq e^{s\varkappa\Gamma^{\star}}e^{-s\Gamma_{l,j}\int\mu'(x)\pi(x)\,dx}.
\]
Furthermore, since $\mu\pi',\mu'\pi\in L^{1}(\mathbb{R})$, we have
\[
\int\mu'(x)\pi(x)\,dx=-\int\mu(x)\pi'(x)\,dx=2\int\mu^{2}(x)\pi(x)\,dx>0
\]
yielding the final bound
\[
\zeta_{l,j}(s) \leq e^{s\kappa \Gamma^*}e^{-2s\Gamma_{l,j}\alpha}
\]
where $\Gamma_{l,j} = \sum\limits_{k=l}^{j}\gamma_k, \alpha = \int \mu^2(x)\pi(x)\,dx$. Since $\sum\limits_{k=1}^{\infty}\gamma_k = \infty$, we may estimate
\begin{align*}
\sum\limits_{p=j+1}^{\infty} \gamma_p e^{-2\alpha \Gamma_{j+1,p}} < \int\limits_{0}^{+\infty}e^{-2\alpha x}\,dx = \frac{1}{2\alpha}
\end{align*}
yielding 
\begin{align*}
\sum_{p=j+1}^{q+1}\gamma_{p} \zeta^{1/4}_{j+1,p}(4) \leq e^{\kappa \Gamma^*}\sum_{p=j+1}^{q+1}\gamma_{p}e^{-2\alpha \Gamma_{j+1,p}} \leq \frac{e^{\kappa \Gamma^*}}{2\alpha} 
\end{align*}
Using the same integral bound, 
\begin{align*}
\mathsf{E}\left[\sum_{j=l}^{q}\gamma_{j}\prod_{k=l}^{j}\alpha_{k}\left(\sum_{p=j+1}^{q+1}\gamma_{p}\prod_{k=j+1}^{p}\alpha_{k}\right)^{2}\right] \leq
\frac{e^{2\kappa \Gamma^*}}{2\alpha}\sum_{j=l}^{q}\gamma_{j}e^{-2\alpha \Gamma_{l,j}} \leq \frac{e^{2\kappa \Gamma^*}}{4\alpha^2}
\end{align*}
Plugging into bound for conditional variance, 
\begin{align*}
\mathsf{Var}\left[\sum_{p=l}^{q}\gamma_{p+1}\partial_{y_{l}}^{K}f\left(X_{p}\right)\Big|X_{l-1}\right] \leq
\frac{A_{K}^{2}e^{2\kappa \Gamma^*}}{4\alpha^2(1-\gamma_1 B_\mu)^3}
\end{align*}
\par
The proof is completed.
\end{proof}

\subsection{Proof of Lemma~\ref{a_decay}}

We have for any \(\mathbf{k}\neq \mathbf{0}\)
\begin{eqnarray*}
a_{p,l,\mathbf{k}}(x)=\mathsf{E}\left[\mathbf{H}_\mathbf{k}\left(Z_l\right)[Q_{p,l}\left(\Phi_l(x,Z_l)\right)-Q_{p,l}\left(\Phi_l(x,0)\right)]\right].
\end{eqnarray*}
Hence
\begin{eqnarray*}
\|a_{p,l,\mathbf{k}}\|_{\infty}\leq \sqrt{\gamma_l}\|\nabla Q_{p,l}\|_{\infty}.
\end{eqnarray*}
Since \(f,\mu\in C^1(\mathbb{R}^d),\)  we have
\begin{eqnarray*}
\nabla Q_{p,l}(x)=\mathsf{E}\left[(I-\gamma_{l+1}\nabla\mu(x))\nabla Q_{p,l+1}(x-\gamma_{l+1}\mu(x)+\sqrt{\gamma_{l+1}}\xi)\right].
\end{eqnarray*}
Hence
\begin{eqnarray*}
\|\nabla Q_{p,l}\|_\infty\leq (1-b_{\mu}\gamma_{l+1})\|\nabla Q_{p,l+1}\|_\infty
\end{eqnarray*}
and
\begin{eqnarray*}
\|\nabla Q_{p,l}\|_\infty&\leq & \|\nabla f\|_\infty\prod_{r=l+1}^p (1-b_{\mu}\gamma_{r})
\\
&\leq & \|\nabla f\|_\infty\exp\left(-b_{\mu}\sum_{r=l+1}^p \gamma_{r}  \right).
\end{eqnarray*}
As a result
\begin{eqnarray*}
\|a_{p,l,\mathbf{k}}\|_{\infty}\leq \sqrt{\gamma_l}\, \|\nabla f\|_\infty\exp\left(-b_{\mu}\sum_{r=l+1}^p \gamma_{r}  \right).
\end{eqnarray*}

\appendix

\section{Bounds for moments of  ULA}\label{sec:appendix}
In this appendix we give some definitions and prove auxiliary results which are the used to bound the exponential moment of the additive functional of ULA chain (Theorem~\ref{th:exp_mom_ula}).
\par
For $\gamma>0$, the Markov kernel $R_\gamma$ on $(\mathbb R^d,\mathcal B(\mathbb R^d))$, corresponding to the Unadjusted Langevin Algorithm with step $\gamma$ is defined by
\begin{equation}
\label{eq:kernel}
R_\gamma(x,A)=\int_{A} \frac{1}{(4 \pi \gamma)^{d/2}} \text{ exp} \left\{ -\frac{1}{4\gamma} \left\| y - x + \gamma \nabla U(x)\right\|^2\right\}dy,\quad A\in \mathcal B(\mathbb R^d),
\end{equation}
while, for $k,l\in\mathbb N$, $k\le l$,
the kernel $Q_{\boldsymbol{\gamma}}^{k,l}$ is
$
Q_{\boldsymbol{\gamma}}^{k,l}=R_{\gamma_l}\cdots R_{\gamma_k}.
$
We make the following assumptions.

\begin{description}
\item[{\bf (H1)}] Assume that $\nabla U$ is Lipschitz, that is, there exists $L\geq 0$ such that 
\begin{eqnarray*}
\|\nabla U(x)-\nabla U(y)\|\leq L\|x-y\|, \quad x,y\in \mathbb{R}^d.
\end{eqnarray*}

\item[{\bf (H2)}] There exist $K>0,$ $M>0$ and $m>0$ such that for any $x\not\in B_K(0)	$ and any $y\in \mathbb{R}^d$ it holds 
\begin{eqnarray*}
\langle D^2U(x)y,y\rangle \geq m\|y\|^2, \quad \|D^3 U(y)\| \leq M. 
\end{eqnarray*}
\end{description}

Due to \cite[Lemma~14]{brosse:2018:diffusion}, conditions {\bf (H1)} and {\bf (H2)} implies that there exist constants $K_1 > 0, m > 0$, such that for any $x \not\in B_{K_1}(0)$ it holds
\begin{equation}
\label{eq:H3}
\langle \nabla U(x), x \rangle \geq \frac{m}{2}\|x\|^2
\end{equation}

We would show that conditions {\bf (H1)} and {\bf (H2)} are sufficient to show that ULA kernel $R_\gamma$ satisfies the so-called drift condition:
\begin{definition}
\label{def:drift}
We say that the Markov kernel $R_\gamma$ satisfies Foster-Lyapunov drift condition if there exist a measurable function $V: X \rightarrow [1; +\infty)$, real numbers $\lambda \in (0,1)$ and $C > 0$ such that for any $x \in X$,
\begin{eqnarray*}
RV(x) \leq \lambda^{\gamma} V(x) + \gamma C
\end{eqnarray*}
\end{definition}

\begin{lem}
\label{lem:drift}
 Assume that the potential $U(x), x \in \mathbb{R}^d$ satisfies conditions {\bf (H1)} and {\bf (H2)} and without loss of generality consider $\nabla U(0) = 0$. Then the kernel $R_\gamma$ from \ref{eq:kernel} satisfies drift condition \ref{def:drift} for any $0 < \gamma < \overline{\gamma} = \frac{m}{4L^2}$ with drift function $V(x) = 1 + \|x\|^2$, constants $\lambda = \exp{\left(-\frac{m}{2}\right)}$, $C = \frac{25K_1^2}{8} + 2d + m$ with $K_1$ from \ref{eq:H3}, $m$ from {\bf (H2)}. 
\end{lem}

\begin{proof} Consider $V(x) = 1 + \|x\|^2$, then
$$
R_{\gamma}V(x) = \int\limits_{\mathbb{R}^d}V(y)P_{\gamma}(x,dy) = \int\limits_{\mathbb{R}^d}(1+\|y\|^2)\frac{1}{(4\pi\gamma)^{\frac{d}{2}}}\exp{\left(-\frac{\|y-x+\gamma\nabla U(x)\|^2}{4\gamma}\right)}\,dy = 
$$

$$
=1 + \frac{1}{(4\pi\gamma)^{\frac{d}{2}}}\int\limits_{\mathbb{R}^d}\|z + x - \gamma \nabla U(x)\|^2 \exp{\left(-\frac{\|z\|^2}{4\gamma}\right)}\,dz
$$
Let us first consider the case $x \notin B(0,K_1)$. Note that 
$$
\|z+x-\gamma \nabla U(x)\|^2 = \|z\|^2 + 2 \langle z, x - \gamma \nabla U(x) \rangle + \gamma^2 \|x - \gamma \nabla U(x)\|^2
$$ 
and the linear term vanishes after integration, moreover, for $Z \sim \mathcal{N}(0,2\gamma I_d)$, it holds $\mathbb{E}\|Z\|^2 = 2\gamma d$. It remains to notice that due to {\bf H1} and {\bf H2} (which implies \ref{eq:H3}),
$$
\|x - \gamma \nabla U(x) \|^2 = \|x\|^2 - 2\gamma \langle \nabla U(x), x\rangle + \gamma^2 \|\nabla U(x)\|^2 \leq
$$

$$
 \leq (1-\gamma m)\|x\|^2 + 2\gamma^2 L^2 \|x\|^2
$$
Thus, plugging everything into expression for $R_{\gamma}V$ and using $\gamma < \frac{m}{4L}$, we obtain
$$
R_{\gamma}V(x) \leq (1-\gamma m + 2\gamma^2 L^2)V(x) + 2\gamma d + (\gamma m - 2\gamma^2 L^2) \leq \exp^{-\frac{\gamma m}{2}}V(x) + \gamma (2d + m)
$$
Now let $x \in B(0,K_1)$. Then simply using $\|x-\gamma \nabla U(x)\|^2 \leq 2(1+L\gamma)^2\|x\|^2$, we obtain
$$
R_{\gamma}V(x) \leq (1-\gamma m + 2\gamma^2 L^2)V(x) + \gamma\left((m - 2\gamma L^2)(1+\|x\|^2) + 2d + 2(1+L\gamma)^2\|x\|^2\right) \leq 
$$

$$
\leq \exp^{-\frac{\gamma m}{2}}V(x) + \gamma(\frac{25K_1^2}{8} + 2d + m)
$$
\end{proof}

For now let us assume for simplicity that we run ULA with fixed step size $\gamma$. Then it is known that under assumption {\bf (H1)} corresponding Markov chain would have unique stationary distribution $\pi_\gamma$, which is different from $\pi$. Yet this chain will be $V-$geometrically ergodic due to \cite[Theorem~19.4.1]{douc:moulines:priouret:soulier:2018}. Namely, the following lemma holds:

\begin{lem}
\label{lem:v_ergodicity}
 Assume that the potential $U(x), x \in \mathbb{R}^d$ satisfies conditions {\bf (H1)} and {\bf (H2)}. Then for $0 < \gamma < \overline{\gamma}=\frac{m}{4L^2}$, for any $x \in X$ it holds
$$
d_V(\delta_xQ_{\boldsymbol{\gamma}}^{1,n},\pi_\gamma) \leq C\rho^n\left(V(x) + \pi_\gamma(V)\right)
$$ 
with $V(x) = 1 + \|x\|^2$ and constants
$$
C = \left(1 + \exp{\left(-\frac{m\gamma}{2}\right)}\right)\left(1+\frac{\overline{b}}{(1-\varepsilon)(1-\exp{\left(-\frac{m\gamma}{2}\right)} - \frac{2b}{1+d})}\right);
$$

$$
b = \gamma(\frac{25K_1^2}{8} + 2d + m); \quad \overline{b} = b\exp^{-\frac{m\gamma}{2}} + d; \quad \varepsilon = 2\Phi\left(-\frac{\sqrt{d}(1+L\gamma)}{2\sqrt{\gamma}}\right); \quad
$$

$$
\rho = \exp{\left(-\gamma\left(\frac{m}{2} - 2\frac{\frac{25K_1^2}{8} + 2d + m}{d}\right)\frac{\log{(1-\varepsilon)}}{\log{(1-\varepsilon)} + \log{(\exp{\left(-\frac{m\gamma}{2}\right)} + \frac{2b}{d+1})}}\right)}
$$
\end{lem}

\begin{proof} Note that the condition {\bf (H1)} implies that the Markov kernel $Q^{1,n}_{\boldsymbol{\gamma}}$ satisfies $(1,\varepsilon)$-Doeblin condition with $\varepsilon = 2\Phi\left(-\frac{\sqrt{d}(1+L\gamma)}{2\sqrt{\gamma}}\right)$. Together with drift condition \ref{eq:drift} it allows to apply \cite[Theorem~19.4.1]{douc:moulines:priouret:soulier:2018} with appropriate constants. 
\end{proof}

Aforementioned lemmas allow us to bound the exponential moment of the additive functional of ULA. Namely, the following theorem holds

\begin{thm}
\label{thm:ula_const_size}
Suppose that ULA kernel \ref{eq:kernel} satisfies assumptions {\bf (H1)} and {\bf (H2)}, and let $X_1,\ldots,X_n,\ldots$ be generated by ULA with constant step size $\gamma$. Let $X_0 = x$ be fixed. Then for any bounded function $g(x) : \mathbb{R}^d \rightarrow \mathbb{R}, |g(x)| \leq M$, it holds
\begin{eqnarray*}
\mathsf{E}\left[\exp\left(-\rho\gamma\sum_{k=l}^{p}\left(g(X_{k})-\pi_{\gamma}(g)\right)\right)\right] \leq C_1\exp{\left(s^2 \kappa \gamma^2 (p-l+1)\right)}
\end{eqnarray*}
for some absolute constant $C_1$ not depending on $l,p$.
\end{thm}

\begin{proof} Conditions {\bf (H1)} and {\bf (H2)} imply that the Markov kernel $R_\gamma$ is $V-$ergodic (lemma \ref{lem:v_ergodicity}) and satisfies drift condition \ref{lem:drift}. Due to \cite[Theorem~3, Fact 3]{moulines:bound_dif:2019}, we obtain the following bound
\begin{align*}
\mathsf{E}_x\left[\exp\left(-s\gamma\sum_{k=l}^{p}\left(g(X_{k}) - \mathsf{E}_x g(X_k)\right)\right)\right] \leq \exp{\left(s^2 \kappa \gamma^2 (p-l+1)\right)}
\end{align*}
\sergey{I will put precise constant later}
with constant $\kappa = ...$. Note that
\begin{align*}
\mathsf{E}\left[\exp\left(-s\gamma\sum_{k=l}^{p}\left(g(X_{k})-\pi_{\gamma}(g)\right)\right)\right] = 
\end{align*}

\begin{align*}
=\mathsf{E}_x\left[\exp{\left(-s\gamma\sum_{k=l}^{p}\left(g(X_{k}) - \mathsf{E}_x g(X_k)\right)\right)}\right]
\exp{\left(s\gamma\sum\limits_{k=l}^{p}\left(\pi_\gamma (g) - \mathsf{E}_x g(X_k)\right)\right)}
\end{align*}
Using lemma~\ref{lem:v_ergodicity},
\begin{align*}
|\mathsf{E} g(X_k) - \pi_\gamma(g)| \leq Md_V(\pi_\gamma, \delta_xQ^{1,n}_{\boldsymbol{\gamma}}) \leq CM\rho^k\left(V(x) + \pi_\gamma(V)\right)
\end{align*}
Hence,
\begin{align*}
\exp{\left(s\gamma\sum\limits_{k=l}^{p}\left(\pi_\gamma (g) - \mathsf{E}_x g(X_k)\right)\right)} \leq \exp{\left(s\gamma CM(V(x) + \pi_\gamma(V))\frac{\rho^l - \rho^{p+1}}{1-\rho}\right)}
\end{align*}
and the statement follows.
\end{proof}

Let us formulate and prove prove analogue of the previous theorem for inhomogeneous Markov chain with step size $\gamma_1,\ldots,\gamma_n,\ldots$. Note that under assumptions $\sum_{p=1}^{\infty} \gamma_p=\infty, \quad \sum_{p=1}^{\infty} \gamma^2_p<\infty$ and {\bf (H1)} the stationary distribution of ULA-based chain will be equal to $\pi$. We will need additional assumption
\begin{description}
\item[{\bf (H3)}] There exist such constants $\rho > 0, K_3 > 0$ that 
\begin{eqnarray*}
d_V(\delta_xQ_{\boldsymbol{\gamma}}^{1,n},\pi) \leq K_3 V(x) \rho^{\Gamma_{1,n}} 
\end{eqnarray*}
\end{description}

\begin{thm}
\label{th:exp_mom_ula}
Suppose that
\begin{eqnarray*}
\sum_{p=1}^{\infty} \gamma_p=\infty, \quad \sum_{p=1}^{\infty} \gamma^2_p<\infty.
\end{eqnarray*}
Under assumptions {\bf (H1)}, {\bf (H2)} and {\bf (H3)}, for any bounded function $g$ on $\mathbb{R}^d$ 
we have
%\begin{eqnarray*}
%\mathsf{E}\left[\left(\sum_{p=l}^{N+n}\omega_{p,n}^{N}g(X_{p})-\bar Q_l(X_l)\right)^{4}\right] \lesssim \frac{1}{\Gamma^4_{N+2,N+n+1}}
%\end{eqnarray*}
 %and
%\begin{eqnarray*}
%\bar Q_l(X_l)=\mathsf{E}\left[\left.\sum_{p=l}^{N+n}\omega_{p,n}^{N}f(X_{p})\right|X_{l}\right].
%\end{eqnarray*}
%Moreover, 
\begin{eqnarray}
\label{eq:exp_decrease_step}
\mathsf{E}\left[\exp\left(-s\left(\sum_{k=l}^{p}\gamma_{k+1}g(X_{k})-\left(\sum_{k=l}^{p}\gamma_{k+1}\right)\pi(g)\right)\right)\right]\leq C
\end{eqnarray}
for all natural $0<l\leq p<\infty$ with constant $C$ not depending on $l,p.$. Moreover, 
\begin{eqnarray}
\label{eq:var_decrease_step}
\mathsf{Var}_x{\left(\sum_{k=l}^{l+n-1}\frac{\gamma_{k+1}}{\Gamma_{l+1,l+n}}g(X_k) - \mathsf{E}_x \sum_{k=l}^{l+n-1}\frac{\gamma_{k+1}}{\Gamma_{l+1,l+n}}g(X_k)\right)} \leq \frac{C_1}{\Gamma_{l+1,l+n}}
\end{eqnarray}
\end{thm}
\begin{proof}

 Note that under assumptions {\bf (H1)}, {\bf (H2)} and {\bf (H3)} it holds due to \cite[Theorem~4]{moulines:bound_dif:2019} that
 \begin{align}
 \label{eq:exp_bounded_dif}
\mathsf{E}_x\left[\exp\left(-s\left(\sum_{k=l}^{p}\gamma_{k+1}g(X_{k}) - \mathsf{E}_x \left(\sum\limits_{k=l}^{p}\gamma_{k+1}g(X_k)\right)\right)\right)\right] \leq \exp{\left(s^2 \kappa \sum\limits_{k=l}^{p}\gamma_{k+1}^2\right)} \leq \exp{\left(s^2 \kappa \Gamma^*\right)}
\end{align}
\sergey{This theorem is not yet written in the second paper}
with $\Gamma^* = \sum\limits_{k=1}^{\infty}\gamma_k^2 < \infty$ and $\kappa = ...$. Using the same transformation as in theorem~\ref{thm:ula_const_size}, and using the bound
\begin{align*}
|\mathsf{E}_x f(X_k) - \pi(f)| \leq K_3 M \rho^{\Gamma_{1,k}}V(x)
\end{align*}
we obtain
\begin{align*}
\mathsf{E}\left[\exp\left(-s\left(\sum_{k=l}^{p}\gamma_{k+1}g(X_{k})-\left(\sum_{k=l}^{p}\gamma_{k+1}\right)\pi(g)\right)\right)\right] \leq \exp{\left(s^2 \kappa \Gamma^*\right)}\exp{\left(K_3 M V(x) \sum\limits_{k=l}^{p}\rho^{\Gamma_{1,k}}\gamma_{k+1}\right)} \leq
\end{align*}

\begin{align*}
\leq \exp{\left(s^2 \kappa \Gamma^*\right)}\exp{\left(K_3 M V(x) \rho^{\Gamma_{1,k}} \int\limits_{0}^{+\infty}\rho^y\,dy\right)} = \exp{\left(s^2 \kappa \Gamma^*\right)}\exp{\left(K_3 M V(x) \frac{\rho^{\Gamma_{1,k}}}{\ln{\rho}}\right)}
\end{align*}
\end{proof}
which proves \ref{eq:exp_decrease_step}. Note that from \ref{eq:exp_bounded_dif} by Taylor expansion for small $s$ we obtain
\begin{align*}
1 + \frac{s^2}{2}\mathsf{Var}_x{\left(\sum_{k=l}^{l+n-1}\gamma_{k+1}g(X_k) - \mathsf{E}_x \sum_{k=l}^{l+n-1}\gamma_{k+1}g(X_k)\right)} + O(s^3) \leq 1 + s^2 \kappa \sum\limits_{k=l+1}^{l+n}\gamma_{k}^2 + O(s^4)
\end{align*}
Since it holds for arbitarily small $s$, the variance is bounded by
\begin{align*}
\mathsf{Var}_x{\left(\sum_{k=l}^{l+n-1}\gamma_{k+1}g(X_k) - \mathsf{E}_x \sum_{k=l}^{l+n-1}\gamma_{k+1}g(X_k)\right)} \leq 2\kappa \sum\limits_{k=l+1}^{l+n}\gamma_{k}^2 \leq 2\kappa \Gamma_{l+1,l+n}
\end{align*}
Now we need to divide both parts by $\Gamma^2_{l+1,l+n}$ to get \ref{eq:var_decrease_step}. The proof is completed.
\bibliographystyle{plain}
\bibliography{refs-1}
\end{document}